---
title: ''
output: html_document
---


## About  

This section has tables and graphs, in separate tabs.  

Anything we definitely want to show up here in the explanation?  


<details>
<summary>Kim's to-dos</summary>

Tasks to address from analysis plan:  

+  QAQC flag summary - make it obvious if there are any 1s or -3s in the dataset  
+  Something to check percent cover standardization (e.g., max. cover).  
+  Remove plots that haven't been sampled at least X # of times (currently X = 3) - but maybe wait until analysis step to do this, so they show up in stacked and spatial bar charts?    
+  Remove flagged data. Make row in input spreadsheet so user can specify which flagged data to remove. Modify function to use user's input (rather than my hard-coded -3)  


Need to do something about NMST abiotic categories, unknown species, unvegetated, etc.  

Probably want data frames by:  

-  species (original)  
-  vegetation group  
-  with custom metrics  


Tasks from analysis plan that have been addressed in this .Rmd:  

+  Identify rows of missing data (plots not measured)  
+  table of plots ~ year  

</details>

## Missing/Removed Data  

### Plots/dates without samples  

There were **`r nrow(unsampds)` rows** with no data. These rows were removed from the dataset before further processing. If rows were removed, relevant information is in a table below.   

```{r}
if(nrow(unsampds) > 0){
    unsampds %>% 
        DT::datatable(rownames = FALSE,
                      caption = "Plots/Dates with no sampling information") 
}
```



### Data flagged suspect or reject  

There were **`r nrow(susps)` data points** that were removed due to QA/QC flags. If data points were removed, relevant information is in a table below.


```{r, message = FALSE}
if(nrow(susps) > 0){
    susps %>% 
        DT::datatable(rownames = FALSE,
                      caption = "Removed data points") 
}
```

.

```{r}
# all commented out 7/3/23 because I don't think we'll use it

# # Joining by species category  
# # species group totals 
# cat_sums <- sum_spp_categories(data = dat)
# 
# # join to sampling info, then pivot longer for easier plotting
# dat_summ <- cbind(dat[, 1:which(names(dat) == "Notes")], cat_sums)
# names(dat_summ)[which(is.na(names(dat_summ)))] <- "Other"
# 
# dat_summ_long <- dat_summ %>% 
#     pivot_longer(-(Reserve:Notes),
#                  names_to = "Group",
#                  values_to = "Cover")


```


## Sampling Info  

### Sites and sample dates  

```{r}
table(dat$StTrnsPlt, dat$Year) %>%
    kable(caption = "# samples at each plot in each year") %>%
    kable_styling("striped") %>%
    scroll_box(width = "600px", height = "300px")
```

### Species  

```{r, message = FALSE}
DT::datatable(spp_summ, 
              rownames = FALSE,
              caption = "Species cover and presence per site, across all years. This table is sortable and searchable.")
```


## Exploratory Graphics   

In the final output, the species highlighted in these graphics will be those specified by each Reserve in their input spreadsheet. For development, I've been using the most common 3 species and grouping everything else into Other.  

The plot below is essentially the 'averaged stacked bar chart' by zone, but *showing each data point* rather than averaged; and not stacked bars.  

### Through time, by zone  

```{r, message = FALSE}
plot_through_time(data = top_3_mean,
          group = Species,
          panels = Vegetation_Zone) +
    scale_color_manual(values = palcols) +
    ggtitle("Top 3 species by mean cover") +
    theme_bw()
```


### Through time, all plots combined   

```{r, message = FALSE}
ggplot(top_3_mean, aes(x = Year, y = Cover, 
                     group = Species, color = Species, fill = Species)) +
        geom_jitter(alpha = 0.2) +
        geom_smooth(method = "loess", se = FALSE) +
    scale_color_manual(values = palcols) +
    ggtitle("Top 3 species by mean cover, all plots combined") +
    theme_bw()
```

### Exploratory - EIR  

```{r}
ggplot(dat) +
    geom_jitter(aes(x = Year, y = EIR, color = Vegetation_Zone),
                alpha = 0.4) +
    facet_wrap(~Vegetation_Zone) +
    theme(legend.position = "none") +
    theme_bw() +
    labs(title = "EIR by Zone",
         subtitle = "as calculated per plot/date combination")
```



```{r}
ggplot(dat) +
    geom_jitter(aes(x = Year, y = EIR, color = SiteID),
                alpha = 0.4) +
    facet_wrap(~Vegetation_Zone) +
    # theme(legend.position = "none") +
    theme_bw() +
    labs(title = "EIR by Site, within Zone")
```


```{r}
ggplot(dat) +
    geom_histogram(aes(x = EIR, fill = Vegetation_Zone),
                   col = "gray40",
                   binwidth = 0.1) +
    facet_wrap(~Vegetation_Zone) +
    labs(title = "Histogram of EIR by Marsh Zone") +
    theme_bw() +
    theme(legend.position = "none")
```





