---
title: ''
output: html_document
---

<details>
<summary>Kim's to-dos</summary>

Tasks to address from analysis plan:  

+  QAQC flag summary - make it obvious if there are any 1s or -3s in the dataset  
+  Something to check percent cover standardization (e.g., max. cover).  
+  Remove plots that haven't been sampled at least X # of times (currently X = 3) - but maybe wait until analysis step to do this, so they show up in stacked and spatial bar charts?    
+  Remove flagged data. Make row in input spreadsheet so user can specify which flagged data to remove. Modify function to use user's input (rather than my hard-coded -3)  


Need to do something about NMST abiotic categories, unknown species, unvegetated, etc.  

Probably want data frames by:  

-  species (original)  
-  vegetation group  
-  with custom metrics  


Tasks from analysis plan that have been addressed in this .Rmd:  

+  Identify rows of missing data (plots not measured)  
+  table of plots ~ year  

</details>

## Tables  

### Plots/dates without samples  

There were **`r nrow(find_unsampleds(dat))` rows** with no data. These rows were removed from the dataset before further processing. If rows were removed, relevant information is in a table below.   

```{r}
if(nrow(find_unsampleds(dat)) > 0){
    find_unsampleds(dat) %>% 
        DT::datatable(rownames = FALSE,
                      caption = "Plots/Dates with no sampling information")
}
```


```{r}
# get rid of unsampled rows and turn remaining NAs into 0s
dat <- remove_unsampleds(dat)  # get rid of rows full of NA
dat <- na_to_0(dat)  # turn remaining NAs into 0s
```


### Data flagged suspect or reject  



```{r, message = FALSE}

dat <- remove_suspect_values(dat, flags = c("-3"))  # remove suspect values

# may need to re-remove rows of NA values, in case removing suspect values turned a whole row into NAs
```




```{r, message = FALSE}
dat <- join_zones(dat, stn_tbl)  # add vegetation zones to each row

dat_long <- dat %>%
    pivot_longer(-(Reserve:Notes),
                 names_to = "Species",
                 values_to = "Cover")
```

.

```{r}
# Joining by species category  
# species group totals 
cat_sums <- sum_spp_categories(data = dat)

# join to sampling info, then pivot longer for easier plotting
dat_summ <- cbind(dat[, 1:which(names(dat) == "Notes")], cat_sums)
names(dat_summ)[which(is.na(names(dat_summ)))] <- "Other"

dat_summ_long <- dat_summ %>% 
    pivot_longer(-(Reserve:Notes),
                 names_to = "Group",
                 values_to = "Cover")


```



```{r, message = FALSE}
# combine species  
top_5_mean <- lump_species(dat_long, "mean", 5)  # species with highest 5 overall means
top_5_nonzero <- lump_species(dat_long, "nonzero", 5)  # species with 5 highest "present"

```


```{r, message = FALSE}
# attach EIR to data frame
dat_long <- dat_long %>% 
    left_join(eis) %>% 
    mutate(Invader = case_when(is.na(Invader) ~ 0,
                               .default = 1))

EIR <- dat_long %>% 
    select(Reserve:PlotID,
           Species,
           Cover,
           Invader) %>% 
    mutate(EI_Cover = Cover * Invader) %>% 
    group_by(Year, Month, Day, Reserve, SiteID, TransectID, PlotID) %>% 
    summarize(EIR = round(sum(EI_Cover) / sum(Cover), 5))
# confirmed at April TWG that denominator is total cover

dat <- left_join(dat, EIR)
```


### Sites and sample dates  

```{r}
table(dat$StTrnsPlt, dat$Year) %>%
    kable(caption = "# samples at each plot in each year") %>%
    kable_styling("striped") %>%
    scroll_box(width = "600px", height = "300px")
```

### Species  

```{r, message = FALSE}
spp_summ <- 
    dat_long %>% 
    group_by(SiteID, Species) %>% 
    summarize(times_present = sum(Cover > 0),
              mean_cover = round(mean(Cover, na.rm = TRUE), 3),
              max_cover = round(max(Cover, na.rm = TRUE), 3)
              ) %>% 
    arrange(Species, SiteID)

# kable(spp_summ) %>% 
#     kable_styling("striped") %>% 
#     scroll_box(width = "500px", height = "300px")

DT::datatable(spp_summ, 
              rownames = FALSE,
              caption = "Species cover and presence per site, across all years. This table is sortable and searchable.")
```


```{r, message = FALSE}
top_3_mean <- lump_species(dat_long, "mean", 3) %>% 
    dplyr::mutate(Date = lubridate::decimal_date(lubridate::ymd(paste(Year, Month, Day, sep = "-"))),
                  Years_sinceStart = round(Date - min(Date), 4),
                  PlotID_full = paste(SiteID, TransectID, PlotID, sep = "-")) %>% 
    dplyr::relocate(c(PlotID_full, Date, Years_sinceStart), .before = Year)


top_3_wide <- top_3_mean %>% 
    pivot_wider(names_from = Species,
                values_from = Cover,
                values_fill = 0) %>% 
    group_by(PlotID_full) %>% 
    mutate()


stn_tbl <- stn_tbl %>% 
    mutate(PlotID_full = paste(SiteID, TransectID, PlotID, sep = "-"))

# establish color palette for all graphics in QA/QC child doc
palcols <- RColorBrewer::brewer.pal(4, "Paired")
names(palcols) <- unique(top_3_mean$Species)
```

## Exploratory Graphics   

In the final output, the species highlighted in these graphics will be those specified by each Reserve in their input spreadsheet. For development, I've been using the most common 3 species and grouping everything else into Other.  

The plot below is essentially the 'averaged stacked bar chart' by zone, but *showing each data point* rather than averaged; and not stacked bars.  

### Through time, by zone  

```{r, message = FALSE}
plot_through_time(data = top_3_mean,
          group = Species,
          panels = Vegetation_Zone) +
    scale_color_manual(values = palcols) +
    ggtitle("Top 3 species by mean cover") +
    theme_bw()
```


### Through time, all plots combined   

```{r, message = FALSE}
ggplot(top_3_mean, aes(x = Year, y = Cover, 
                     group = Species, color = Species, fill = Species)) +
        geom_jitter(alpha = 0.2) +
        geom_smooth(method = "loess", se = FALSE) +
    scale_color_manual(values = palcols) +
    ggtitle("Top 3 species by mean cover, all plots combined") +
    theme_bw()
```

## Exploratory - EIR  

```{r}
ggplot(dat) +
    geom_jitter(aes(x = Year, y = EIR, color = Vegetation_Zone),
                alpha = 0.4) +
    facet_wrap(~Vegetation_Zone) +
    theme(legend.position = "none") +
    theme_bw() +
    labs(title = "EIR by Zone",
         subtitle = "as calculated per plot/date combination")
```



```{r}
ggplot(dat) +
    geom_jitter(aes(x = Year, y = EIR, color = SiteID),
                alpha = 0.4) +
    facet_wrap(~Vegetation_Zone) +
    # theme(legend.position = "none") +
    theme_bw() +
    labs(title = "EIR by Site, within Zone")
```


```{r}
ggplot(dat) +
    geom_histogram(aes(x = EIR, fill = Vegetation_Zone),
                   col = "gray40",
                   binwidth = 0.1) +
    facet_wrap(~Vegetation_Zone) +
    labs(title = "Histogram of EIR by Marsh Zone") +
    theme_bw() +
    theme(legend.position = "none")
```





