---
title: ''
output: html_document
---

## About  

This section has summary graphics. The color palette is meant to be colorblind-friendly but please reach out if you are colorblind and it's not working for you.  


## Averaged Stacked Bar Charts  


### By Vegetation Zone  

```{r, message = FALSE}
top_3_mean %>% 
    group_by(Vegetation_Zone, Year, Species) %>% 
    summarize(Cover = mean(Cover, na.rm = TRUE)) %>% 
    ggplot(aes(x = Year, y = Cover,
               group = Species, fill = Species),
           alpha = 0.8) +
    geom_col(position = "stack") +
    facet_wrap(~Vegetation_Zone) +
    scale_fill_manual(values = palcols) +
    theme_bw() +
    labs(title = "Vegetation Cover through time",
         subtitle = "averaged by Species within Zone")
```


### By Site  

```{r, message = FALSE}
top_3_mean %>% 
    group_by(SiteID, Year, Species) %>% 
    summarize(Cover = mean(Cover, na.rm = TRUE)) %>% 
    ggplot(aes(x = Year, y = Cover,
               group = Species, fill = Species),
           alpha = 0.8) +
    geom_col(position = "stack") +
    facet_wrap(~SiteID) +
    scale_fill_manual(values = palcols) +
    theme_bw() +
    labs(title = "Vegetation Cover through time",
         subtitle = "averaged by Species within Site")
```


### By Zone within Site  

```{r, message = FALSE}
top_3_mean %>% 
    group_by(SiteID, Vegetation_Zone, Year, Species) %>% 
    summarize(Cover = mean(Cover, na.rm = TRUE)) %>% 
    ggplot(aes(x = Year, y = Cover,
               group = Species, fill = Species),
           alpha = 0.8) +
    geom_col(position = "stack") +
    facet_grid(Vegetation_Zone~SiteID) +
    scale_fill_manual(values = palcols) +
    theme_bw() +
    labs(title = "Vegetation Cover through time",
         subtitle = "averaged by Species within Site/Zone combination")
```




## Spatial Bar Charts  

One plot per Site; showing 4 evenly spaced-through-time samples.   

<details>
<summary>Thoughts for further development</summary>

-  Will need to figure out how to do this based on species the reserve chose  
-  was able to generate legend via `ggplot2`, using a colorblind-friendly palette from `RColorBrewer` ("Paired").  
-  for some reserves, these might be better by transect rather than by site (e.g. CBV; at least a couple of sites at GTM)  


Can potentially do this by transect within station as well, and make it a user option in the input spreadsheet. If it's selected, then there will be an additional for loop between the `st in SiteID` and `plt in PlotID` ones, for `tr in TransectID`. For each site -- for each transect -- make base plot, based on coords in transect -- loop through plots to make the subplots and add them to the base plot -- after looping, name base plot based on transect -- use `{patchwork}` to paste together transects within the site so they're all in a row or panel or whatever.    

</details>  

```{r}
# THIS CHUNK IS AN ADAPTATION OF BROOK'S CODE

# for(st in unique(stn_tbl$SiteID)){
#     
#     # subset to the site
#     stn_tmp <- stn_tbl[stn_tbl$SiteID == st, ]
#     
#     long_offset <- (max(stn_tmp$Long) - min(stn_tmp$Long)) * 0.05
#     lat_offset <- (max(stn_tmp$Lat) - min(stn_tmp$Lat)) * 0.125
#     
#     # blank plot canvas
#     plot(c(min(stn_tmp$Long) - long_offset,
#            max(stn_tmp$Long) + long_offset),
#          c(min(stn_tmp$Lat) - lat_offset,
#            max(stn_tmp$Lat) + lat_offset),
#          type = "n",
#          bty = "n",
#          xlab = "Longitude", ylab = "Latitude",
#          main = paste(params$file_code, st))
#     
#     # subset the lumped data to the single site
#     dat_tmp <- top_3_wide[top_3_wide$SiteID == st, ]
#     
#  
# 
#     # loop through each plot at the site
#     # make a barplot of the top 3 species + other
#     # and add as subplot to the blank canvas, based on lat/long
#     for (plt in unique(dat_tmp$PlotID_full)){
#         
#         dat_plt <- dat_tmp[dat_tmp$PlotID_full==plt, ]
#         yrs <- dat_plt$Years_sinceStart
#             
#         # subset the years: min, max, and as close as possible to the difference/3  
# 
#         un_yrs <- unique(yrs)
#         div <- (max(un_yrs) - min(un_yrs)) / 3
#         divs_to_keep <- c(min(un_yrs) + div,
#                           max(un_yrs) - div)
#         
# 
#         # figure out what to keep by finding the closest actual measurement to each, using absolute values
#         keeps2 <- un_yrs[which(abs(un_yrs - divs_to_keep[1]) == min(abs(un_yrs - divs_to_keep[1])))]
#         keeps3 <- un_yrs[which(abs(un_yrs - divs_to_keep[2]) == min(abs(un_yrs - divs_to_keep[2])))]
# 
#         keeps <- c(min(un_yrs), keeps2, keeps3, max(un_yrs))
#         
#         # re-subset the data frame to only those years
#         dat_plt <- dat_plt[dat_plt$Years_sinceStart %in% keeps, ]
#         yrs <- dat_plt$Years_sinceStart
#         
#         lat <- stn_tmp[stn_tmp$PlotID_full==plt, ]$Lat
#         long <- stn_tmp[stn_tmp$PlotID_full==plt, ]$Long
#         subplot(barplot( t(as.matrix(dat_plt[,(ncol(dat_plt)-3):ncol(dat_plt)])),  # last 4 columns are the top 3 species + Other
#                          xlab = "",
#                          axes=FALSE),
#                 x = c(long - long_offset, long + long_offset),
#                 y = c(lat - lat_offset, lat + lat_offset))  
#     }
# }
```





```{r}
# p1 + annotation_custom(ggplotGrob(p2), xmin = 1, xmax = 3, 
#                        ymin = -0.3, ymax = 0.6)

# Tutorial on which the subplot idea below is based:
# https://www.r-bloggers.com/2019/02/plots-within-plots-with-ggplot2-and-ggmap/  
```



```{r, fig.width = 8, fig.height = 4}

# add species into station table for the sole purpose of generating a legend.  
stn_tbl$Species <- rep_len(unique(top_3_mean$Species), length.out = nrow(stn_tbl))

for(st in unique(stn_tbl$SiteID)){
    
    # subset to the site
    stn_tmp <- stn_tbl[stn_tbl$SiteID == st, ]
    
    long_offset <- (max(stn_tmp$Long) - min(stn_tmp$Long)) * 0.05
    lat_offset <- (max(stn_tmp$Lat) - min(stn_tmp$Lat)) * 0.11
    
    # blank plot canvas
    base <- ggplot(stn_tmp, aes(x = Long, y = Lat, 
                                group = Species,
                                fill = Species)) +
        geom_col(na.rm = TRUE) +  # necessary to make legend show up
        labs(title = paste(params$file_code, st),
             x = "Longitude",
             y = "Latitude") +
        scale_x_continuous(limits = c(min(stn_tmp$Long) - long_offset, max(stn_tmp$Long) + long_offset)) +
        scale_y_continuous(limits = c(min(stn_tmp$Lat) - lat_offset, max(stn_tmp$Lat) + lat_offset)) +
        scale_fill_manual(values = palcols) +
        theme_bw() +
        theme(legend.position = "right")
    
    
    # subset the lumped data to the single site
    dat_tmp <- top_3_mean[top_3_mean$SiteID == st, ]
    
 

    # loop through each plot at the site
    # make a barplot of the top 3 species + other
    # and add as subplot to the blank canvas, based on lat/long
    for (plt in unique(dat_tmp$PlotID_full)){
        
        dat_plt <- dat_tmp[dat_tmp$PlotID_full==plt, ]
        yrs <- dat_plt$Years_sinceStart
            
        # subset the years: min, max, and as close as possible to the difference/3  

        un_yrs <- unique(yrs)
        div <- (max(un_yrs) - min(un_yrs)) / 3
        divs_to_keep <- c(min(un_yrs) + div,
                          max(un_yrs) - div)
        

        # figure out what to keep by finding the closest actual measurement to each, using absolute values
        keeps2 <- un_yrs[which(abs(un_yrs - divs_to_keep[1]) == min(abs(un_yrs - divs_to_keep[1])))]
        keeps3 <- un_yrs[which(abs(un_yrs - divs_to_keep[2]) == min(abs(un_yrs - divs_to_keep[2])))]

        keeps <- c(min(un_yrs), keeps2, keeps3, max(un_yrs))
        
        # re-subset the data frame to only those years
        dat_plt <- dat_plt[dat_plt$Years_sinceStart %in% keeps, ]

        # make the subplot
        subplo <- ggplotGrob(ggplot(dat_plt) +
                                 geom_col(aes(x = as.factor(Years_sinceStart),
                                              y = Cover,
                                              fill = Species),
                                          # width = 0.8,
                                          position = "stack",
                                          alpha = 0.8,
                                          show.legend = FALSE) +
                                 scale_fill_manual(values = palcols) +
                                 theme_void() + 
                                 theme(panel.border = element_rect(linetype = "dashed", fill = NA)))
        
        # get coordinates for subplot
        lat <- stn_tmp[stn_tmp$PlotID_full==plt, ]$Lat
        long <- stn_tmp[stn_tmp$PlotID_full==plt, ]$Long
        
        # add the subplot to the graph
        
        base <- base + 
            annotation_custom(grob = subplo,
                              xmin = long - long_offset,
                              xmax = long + long_offset,
                              ymin = lat - lat_offset,
                              ymax = lat + lat_offset)
         
    }
    print(base)
}
```