---
title: ''
output: html_document
---

## About  

Each of the (up to) 4 response variables identified in the Analysis Specs spreadsheet, plus EIR, species richness, and Shannon-Weiner Diversity Index, will have its results appear in its own tab. You will see the same tables and graphics for each.  

### Questions addressed  

-  Are there shifts in key vegetation species/groups over time?  
-  Do these shifts vary by vegetation zone?  
-  Follow-up: Where is the plant community changing (which plots) and what characteristics do those areas have in common (e.g. site, zone, distance from water, elevation)?  

### Reserve-identified response variables:  

The reserve-identified univariate response variables are:  

```{r show-userChoice-univar, results = 'asis'}
cat(paste("*", anaSpec_list$univar$Choice), sep="\n")
```

```{r mod-univar-df}
# widen out the univar data frame
# it's grouped appropriately for these analyses  
dat_univar <- anaSpec_dfs$univar %>% 
    pivot_wider(names_from = Species_or_Group,
                values_from = Cover)

# join EIR - use a join because might not be in same order
EIR_df <- dat %>% 
    select(StTrnsPlt, Year, Month, Day, EIR)

dat_univar <- left_join(dat_univar, EIR_df) %>% 
    left_join(dat_div)

# can join any other metrics here - e.g. richness, diversity  
```

```{r}
# remove plots that don't have at readings in at least 3 separate years
dat_univar <- filter(dat_univar, !(StTrnsPlt %in% not_enough$StTrnsPlt))
```


### Statistical Model  

The main statistical model used is a linear mixed model, via `lme4::lmer()`, with the form `y ~ Vegetation Zone + Time + Zone*Time + 1|Plot`, where `y` is the response variable given in the tab, and each individual plot has a random intercept. 

### Model Diagnostics  

Something here about the section about models throwing errors, and how to tell which one(s) did. Also describe what a singular fit is - when there's not enough variation in the random components to adequately estimate the random effects.  


### Representation of Results  

Coefficient table  

ANOVA table; whether interaction between time and zone is significant  

Estimated marginal slopes (change within each zone)  

Contrasts plot  

```{r}
# set up a list of univariate output info
# estimates, CIs, etc.
univar_outputs <- list()
```



## EIR   

Explanation of what EIR is and how it was calculated.  

Species considered to be 'invasive' within each zone are denoted by 'x' in the following table:  

```{r show-eir-spps}
kableExtra::kable(eir_spps, align = "c") %>% 
    kable_styling("striped") %>%
    scroll_box(width = "800px", height = "200px") 
```


```{r define-univar-EIR}
respn <- "EIR"
dat_univar$resp <- unlist(dat_univar[which(names(dat_univar) == respn)])
```


```{r source-003-EIR, child = "003a_model_run.Rmd"}

```



## Richness  

Species Richness, calculated on all non-abiotic, non-dead, and non-overstory columns; using the `vegan` package.

```{r define-univar-richness}
respn <- "Richness"
dat_univar$resp <- unlist(dat_univar[which(names(dat_univar) == respn)])
```

```{r source-003a-userChosen1b, child = "003a_model_run.Rmd"}

```



## Diversity  

Shannon-Weiner Diversity index, calculated on all non-abiotic, non-dead, and non-overstory columns; using the `vegan` package.  

```{r define-univar-diversity}
respn <- "SWdiv"
dat_univar$resp <- unlist(dat_univar[which(names(dat_univar) == respn)])
```

```{r source-003a-userChosen1c, child = "003a_model_run.Rmd"}

```




```{r, results = "asis"}
respns <- anaSpec_list$univar$Choice
for(i in seq_along(respns)){
    respn <- respns[i]
    cat("\n")
    cat("##", respn, "\n")
    dat_univar$resp <- unlist(dat_univar[which(names(dat_univar) == respn)])
    chld <- knitr::knit_child("003a_model_run.Rmd", quiet = TRUE)
    cat(chld, sep = "\n")
}
```

