---
title: ''
output: html_document
---

## About the Univariate Analyses  

Each of the (up to) 4 response variables identified in the Analysis Specs spreadsheet, and EIR, will have its results appear in its own tab. You will see the same tables and graphics for each.  

<details>
<summary>Kim to-dos and notes</summary>  

List parameters of interest, identified in spreadsheet input.  


For each parameter of interest, run the model, then display tables and figures of model output (including estimated marginal slopes).  


My original version of this used the wide data file, but I think I'm going to be preferring the long one, and filtering based on the selected response parameter.  

</details>


## 1: EIR   

Explanation of what EIR is and how it was calculated.  

Insert table of reserve-identified ecotone invaders, by zone.  

```{r}
respn <- "EIR"
dat$resp <- unlist(dat[which(names(dat) == respn)])
```


```{r, child = "003a_model_run.Rmd"}

```


```{r}
respn <- "Spartina alterniflora"
dat$resp <- unlist(dat[which(names(dat) == respn)])
# but really want to make that adapt to user input
```

## 2: `r respn`  

```{r, child = "003a_model_run.Rmd"}

```


```{r}
respn <- "Juncus roemerianus"
dat$resp <- unlist(dat[which(names(dat) == respn)])
# but really want to make that adapt to user input
```

## 3: `r respn`  

```{r, child = "003a_model_run.Rmd"}

```



```{r}
respn <- "Spartina alterniflora"
dat$resp <- unlist(dat[which(names(dat) == respn)])
# but really want to make that adapt to user input
```


## 4: `r respn`  

```{r, child = "003a_model_run.Rmd"}

```