---
title: ''
output: html_document
---

## About  

Each of the (up to) 4 response variables identified in the Analysis Specs spreadsheet, plus EIR, will have its results appear in its own tab. You will see the same tables and graphics for each.  

The reserve-identified response variables are:  

```{r, results = 'asis'}
cat(paste("*", anaSpec_list$univar$Choice), sep="\n")
```


### Statistical Model  

The main statistical model used is a linear mixed model, via `lme4::lmer()`, with the form `y ~ Vegetation Zone + Time + Zone*Time + 1|Site/Transect/Plot`, where `y` is the response variable given in the tab. For various reasons this might not work as well as we'd hoped, so less complicated models have also been run.  

<details><summary>Expand info on less complicated models</summary>

`y ~ Vegetation Zone + Time + Zone*Time + 1|Transect/Plot` - particularly for reserves with only a single site  

`y ~ Vegetation Zone + Time + Zone*Time + 1|Site/Plot`

`y ~ Vegetation Zone + Time + Zone*Time + 1|Plot` - simplest model, with a random effect for each plot; should work for all reserves but may not account for spatial autocorrelation.  

</details>


### Model Diagnostics  

Something here about the section about models throwing errors, and how to tell which one(s) did. Also describe what a singular fit is - when there's not enough variation in the random components to adequately estimate the random effects.  


### Representation of Results  

Coefficient table  

ANOVA table; whether interaction between time and zone is significant  

Estimated marginal slopes (change within each zone)  

Contrasts plot  



<details>
<summary>Kim to-dos and notes</summary>  

List parameters of interest, identified in spreadsheet input.  


For each parameter of interest, run the model, then display tables and figures of model output (including estimated marginal slopes).  


My original version of this used the wide data file, but I think I'm going to be preferring the long one, and filtering based on the selected response parameter.  

</details>


## 1: EIR   

Explanation of what EIR is and how it was calculated.  

Insert table of reserve-identified ecotone invaders, by zone.  

```{r}
respn <- "EIR"
dat$resp <- unlist(dat[which(names(dat) == respn)])
```


```{r, child = "003a_model_run.Rmd"}

```



```{r}
respn <- anaSpec_list$univar$Choice[1]
# respn <- "Spartina alterniflora"
dat$resp <- unlist(dat[which(names(dat) == respn)])
```

## 2: `r respn`  

```{r, child = "003a_model_run.Rmd"}

```



```{r}
respn <- anaSpec_list$univar$Choice[2]
# respn <- "Juncus roemerianus"
dat$resp <- unlist(dat[which(names(dat) == respn)])
```

## 3: `r respn`  

```{r, child = "003a_model_run.Rmd"}

```




```{r}
respn <- anaSpec_list$univar$Choice[3]
# respn <- "Spartina alterniflora"
dat$resp <- unlist(dat[which(names(dat) == respn)])
```


## 4: `r respn`  

```{r, child = "003a_model_run.Rmd"}

```



```{r}
respn <- anaSpec_list$univar$Choice[4]

if(is.na(respn)){
    respn <- unique(top_3_mean$Species)[1]
}

dat$resp <- unlist(dat[which(names(dat) == respn)])
```


## 5: `r respn`  

```{r, child = "003a_model_run.Rmd"}

```