---
output: html_document
---


```{r}
mod_nested_full <- lmer(resp ~ Years_sinceStart*Vegetation_Zone + (1|SiteID/TransectID/PlotID),
                   data = dat)

mod_nested_trns <- lmer(resp ~ Years_sinceStart*Vegetation_Zone + (1|StTrns/PlotID),
                   data = dat)

mod_plts <- lmer(resp ~ Years_sinceStart*Vegetation_Zone + (1|StTrnsPlt),
                   data = dat)
```


### Models that didn't run  

this code definitely needs to be turned into a function or a loop  
preferable with informative error messages for the end user  
(they won't understand this terminology but for now it's what I'm trying to work out)  

```{r}
# allow for models that have errored and 
# add models that did work to a list to pull out results

mods <- list()


if(exists("mod_nested_full")){
    mods[["fully nested"]] <- mod_nested_full
} else {
    cat("fully nested model errored")
}


if(exists("mod_nested_trns")){
    mods[["transect/plot"]] <- mod_nested_trns
} else {
    cat("transect/plot model errored")
}


if(exists("mod_plts")){
    mods[["plots only"]] <- mod_plts
} else {
    cat("plots only random effect model errored")
}
```


```{r}
trends <- purrr::map(mods, function(x) get_trends(x))
trends_df <- bind_rows(trends, .id = "model") %>% 
    mutate(model = forcats::as_factor(model))
```

Pull out marginal and conditional R^2 for mixed models and add columns to df  

```{r, warning = FALSE, message = FALSE}
rsqds <- purrr::map(mods, function(x) as.data.frame(MuMIn::r.squaredGLMM(x)))
rsqds_df <- bind_rows(rsqds, .id = "model") %>% 
    mutate(model = forcats::as_factor(model))
trends_df <- left_join(trends_df, rsqds_df)
```



### Singular Models  

```{r}
if(exists("mod_nested_full")){check_singularity(mod_nested_full)}
if(exists("mod_nested_trns")){check_singularity(mod_nested_trns)}
if(exists("mod_plts")){check_singularity(mod_plts)}
```

### Table  


```{r}
to_print <- trends_df %>% 
    mutate(across(Years_sinceStart.trend:R2c, function(x) round(x, 4)),
           df = round(df, 2)) %>% 
    arrange(Vegetation_Zone, model) %>% 
    rename(TrendPerYear = Years_sinceStart.trend) %>% 
    relocate(Vegetation_Zone)
    
kbl(to_print) %>% 
    kable_styling("striped") %>%
    scroll_box(width = "600px", height = "300px") 
```


### Numerical effects  

For the model with transect/plot nesting.

```{r}
if(exists("mod_nested_trns")){
    mod_anov <- anova(mod_nested_trns, ddf = "Kenward-Roger")
    tidied <- broom::tidy(mod_anov)
    sig_int <- tidied[tidied$term == "Years_sinceStart:Vegetation_Zone", "p.value"] < 0.05
    print(mod_anov)
    cat("\n \n")
    emtrends(mod_nested_trns, 
                  pairwise ~ Vegetation_Zone, 
                  var = "Years_sinceStart")$emtrends
}
```


### Graphs  

```{r, warning = FALSE, message = FALSE}
if(exists("mod_nested_trns")){
    pdf <- emmip(mod_nested_trns, Vegetation_Zone ~ Years_sinceStart,
                 cov.reduce = range,
                 linearg = list(size = 1), CIs = TRUE,
                 plotit = FALSE)
    
    p1 <- ggplot() +
        geom_line(data = pdf, aes(x = Years_sinceStart,
                                  y = yvar,
                                  col = Vegetation_Zone,
                                  linetype = Vegetation_Zone),
                  linewidth = 0.8) +
        labs(x = "Years elapsed", 
             y = paste("Predicted", respn)) +
        theme_bw()
    
    
    
    p2 <- ggplot() +
        geom_jitter(data = dat, aes(x = Years_sinceStart, y = resp,
                                     col = Vegetation_Zone),
                    alpha = 0.7) +
        geom_line(data = pdf, aes(x = Years_sinceStart,
                                  y = yvar,
                                  col = Vegetation_Zone,
                                  linetype = Vegetation_Zone),
                  linewidth = 0.8) +
        facet_wrap(~Vegetation_Zone) +
        theme_bw() +
        labs(x = "Years elapsed",
             y = respn) +
        theme(legend.position = "none")
    
    print(p1 + ggtitle("Mixed Model fit"))
    print(p2)
}
```


### Contrasts plot  

For the model with transect/plot nesting. If the interaction between vegetation zone and time was significant (p < 0.05), letters are used to represent groups of slopes that are not significantly different (via pairwise comparisons). If the interaction was not significant, this plot represents the marginal slopes but no letters are used.

Some notes from the function. Kenward-Roger method used to estimate denominator degrees of freedom. Confidence level used: 0.95. p-value adjustment: Tukey. alpha = 0.05.  

"If two or more means share the same grouping symbol,
      then we cannot show them to be different.
      But we also did not show them to be the same."

```{r}
if(exists("mod_nested_trns")){
    trnds <- emtrends(mod_nested_trns, 
                  pairwise ~ Vegetation_Zone, 
                  var = "Years_sinceStart")
    mod_cld <- cld(trnds, Letters = letters) %>% 
        mutate(.group = str_trim(.group))
    
    p <- ggplot(mod_cld, aes(x = reorder(Vegetation_Zone, Years_sinceStart.trend),
                             y = Years_sinceStart.trend,
                             label = .group,
                             color = Vegetation_Zone)) +
        geom_hline(yintercept = 0) +
        geom_pointrange(aes(ymin = lower.CL, 
                            ymax = upper.CL),
                        linewidth = 0.8) +
        labs(title = paste("Rate of change in", respn, "by zone"),
             x = "Vegetation Zone",
             y = "change per year") +
        theme_bw() +
        theme(legend.position = "none")
    
    if(sig_int){
        p <- p +
            geom_text(aes(y = max(mod_cld$upper.CL)),
                      nudge_y = max(mod_cld$upper.CL)/10,
                      color = "black")
    }
    print(p)
}
```