---
output: html_document
---


```{r}
mod_plts <- lmer(resp ~ Years_sinceStart*Vegetation_Zone + (1|StTrnsPlt),
                   data = dat_univar)
```


#### Model errors, if any, will show up here  


```{r}
if(exists("mod_plts", inherits = FALSE)){
    cat("\n")
} else {
    cat("plots only random effect model errored")
}
```

```{r}
trends_df <- get_trends(mod_plts)
```

```{r, warning = FALSE, message = FALSE}
# Pull out marginal and conditional R^2 for mixed models and add columns to df  

rsqds <- as.data.frame(MuMIn::r.squaredGLMM(mod_plts))

# trends_df <- left_join(trends_df, rsqds)
```



### Numerical effects  

```{r}
mod_anov <- anova(mod_plts, ddf = "Kenward-Roger")
tidied <- broom::tidy(mod_anov)
sig_int <- tidied[tidied$term == "Years_sinceStart:Vegetation_Zone", "p.value"] < 0.05
mod_trnds <- data.frame(emtrends(mod_plts, 
                                 pairwise ~ Vegetation_Zone, 
                                 var = "Years_sinceStart")$emtrends)




kbl_nmst(mod_anov, 
         caption = paste(respn, " - Type III Analysis of Variance Table, using Kenward-Rogers method"),
         digits = c(4, 4, 1, 2, 2, 4))  
```

```{r}
cat("\n \n \n")

kbl_nmst(rsqds,
         caption = paste(respn, " - Marginal and Conditional R^2 values"),
         digits = 3)
```

```{r}
cat("\n \n \n")

kbl_nmst(mod_trnds, 
         digits = c(4, 4, 4, 2, 4, 4), 
         col.names = c("Veg Zone", "Trend (per year)",
                       "SE", "df", "CI lower", "CI upper"),
         caption = paste(respn, " - Estimated marginal slopes (change per year for each zone)")) 

univar_outputs[[respn]] <- mod_trnds
```


### Graphs  

```{r, warning = FALSE, message = FALSE}
pdf <- emmip(mod_plts, Vegetation_Zone ~ Years_sinceStart,
             cov.reduce = range,
             linearg = list(size = 1), CIs = TRUE,
             plotit = FALSE)

p1 <- ggplot() +
    geom_line(data = pdf, aes(x = Years_sinceStart,
                              y = yvar,
                              col = Vegetation_Zone,
                              linetype = Vegetation_Zone),
              linewidth = 0.8) +
    scale_color_manual(values = palcols_zones) +
    labs(title = paste("Model Predicted Values:", respn),
         x = "Years elapsed", 
         y = paste("Predicted", respn),
         col = "Vegetation Zone",
         linetype = "Vegetation Zone") +
    theme_bw()



p2 <- ggplot() +
    geom_jitter(data = dat_univar, aes(x = Years_sinceStart, y = resp,
                                       col = Vegetation_Zone),
                alpha = 0.7) +
    geom_line(data = pdf, aes(x = Years_sinceStart,
                              y = yvar,
                              col = Vegetation_Zone,
                              linetype = Vegetation_Zone),
              linewidth = 0.8) +
    facet_wrap(~Vegetation_Zone) +
    scale_color_manual(values = palcols_zones) +
    theme_bw() +
    labs(x = "Years elapsed",
         y = respn) +
    theme(legend.position = "none")

print(p1)
# print(p2)


model <- mod_plts
predicted <- estimate_expectation(model, data = "grid")
p3 <- plot(predicted, ribbon = list(alpha = 0.2),
           point = list(alpha = 0.5),
           line = list(alpha = 0.9, size = 0.8)) + 
    theme_bw() +
    facet_wrap(~Vegetation_Zone) + 
    scale_color_manual(values = palcols_zones) +
    labs(title = paste("Predicted response:", respn),
         subtitle = "predictions and CIs generated by easystats packages",
         x = "Years elapsed",
         y = respn) +
    theme(legend.position = "none")

print(p3)
```


### Contrasts plot  

If the interaction between vegetation zone and time was significant (p < 0.05), letters are used to represent groups of slopes that are not significantly different (via pairwise comparisons). If the interaction was not significant, this plot represents the marginal slopes but no letters are used.

Some notes from the function. Kenward-Roger method used to estimate denominator degrees of freedom. Confidence level used: 0.95. p-value adjustment: Tukey. alpha = 0.05.  

"If two or more means share the same grouping symbol,
      then we cannot show them to be different.
      But we also did not show them to be the same."

```{r, fig.height = 5, fig.width = 6}
trnds <- emtrends(mod_plts, 
                  pairwise ~ Vegetation_Zone, 
                  var = "Years_sinceStart")
mod_cld <- cld(trnds, Letters = letters) %>% 
    mutate(.group = str_trim(.group))

mod_cld$Vegetation_Zone <- factor(mod_cld$Vegetation_Zone, levels = zone_factor)

p <- ggplot(mod_cld, aes(x = reorder(Vegetation_Zone, Years_sinceStart.trend),
                         y = Years_sinceStart.trend,
                         label = .group,
                         color = Vegetation_Zone)) +
    geom_hline(yintercept = 0) +
    geom_pointrange(aes(ymin = lower.CL, 
                        ymax = upper.CL),
                    linewidth = 0.8) +
    scale_color_manual(values = palcols_zones) +
    labs(title = paste("Rate of change in", respn, "by zone"),
         x = "Vegetation Zone, low to high rate of change",
         y = "change per year") +
    theme_bw() +
    theme(legend.position = "none")

if(sig_int){
    p <- p +
        geom_text(aes(y = max(mod_cld$upper.CL)),
                  nudge_y = max(mod_cld$upper.CL)/10,
                  color = "black")
}
print(p)
```

Same plot, but x-axis goes from lowest to highest zone.    

```{r, fig.height = 5, fig.width = 6}
p <- ggplot(mod_cld, aes(x = Vegetation_Zone,
                         y = Years_sinceStart.trend,
                         label = .group,
                         color = Vegetation_Zone)) +
    geom_hline(yintercept = 0) +
    geom_pointrange(aes(ymin = lower.CL, 
                        ymax = upper.CL),
                    linewidth = 0.8) +
    scale_color_manual(values = palcols_zones) +
    labs(title = paste("Rate of change in", respn, "by zone"),
         x = "Vegetation Zone, low to high elevation",
         y = "change per year") +
    theme_bw() +
    theme(legend.position = "none")

if(sig_int){
    p <- p +
        geom_text(aes(y = max(mod_cld$upper.CL)),
                  nudge_y = max(mod_cld$upper.CL)/10,
                  color = "black")
}
print(p)
```



```{r}
if(exists("mod_plts")){check_singularity(mod_plts)}
```


```{r}
# cleanup before next response var
rm(p, p1, p2, p3,
   mod_plts,
   trends, trends_df,
   rsqds, rsqds_df, 
   to_print,
   trnds, mod_cld,
   mod_anov, tidied, sig_int, pdf)
```

