---
output: html_document
---

```{r}
mod_plts <- lmer(resp ~ Years_sinceStart*Vegetation_Zone + (1|StTrnsPlt),
                   data = dat_univar)

modyes <- exists("mod_plts", inherits = FALSE)
```

```{r, warning = FALSE, error = FALSE, results = "asis"}
if(exists("mod_plts", inherits = FALSE)){
    # if the model ran, pull out all the info to report in the sections below
    cat("\n")
    
    # tables
    trends_df <- get_trends(mod_plts)
    rsqds <- as.data.frame(MuMIn::r.squaredGLMM(mod_plts))
    
    mod_anov <- anova(mod_plts, ddf = "Kenward-Roger")
    tidied <- broom::tidy(mod_anov)
    sig_int <- tidied[tidied$term == "Years_sinceStart:Vegetation_Zone", "p.value"] < 0.05
    mod_trnds <- data.frame(emtrends(mod_plts, 
                                     pairwise ~ Vegetation_Zone, 
                                     var = "Years_sinceStart")$emtrends)
    
    # save outputs
    univar_outputs[[respn]] <- mod_trnds
    
    # plots to print
    pdf <- emmip(mod_plts, Vegetation_Zone ~ Years_sinceStart,
                 cov.reduce = range,
                 linearg = list(size = 1), CIs = TRUE,
                 plotit = FALSE)
    
    p1 <- ggplot() +
        geom_line(data = pdf, aes(x = Years_sinceStart,
                                  y = yvar,
                                  col = Vegetation_Zone,
                                  linetype = Vegetation_Zone),
                  linewidth = 0.8) +
        scale_color_manual(values = palcols_zones) +
        labs(title = paste("Model Predicted Values:", respn),
             x = "Years elapsed", 
             y = paste("Predicted", respn),
             col = "Vegetation Zone",
             linetype = "Vegetation Zone") +
        theme_bw()
    
    
    
    p2 <- ggplot() +
        geom_jitter(data = dat_univar, aes(x = Years_sinceStart, y = resp,
                                           col = Vegetation_Zone),
                    alpha = 0.7) +
        geom_line(data = pdf, aes(x = Years_sinceStart,
                                  y = yvar,
                                  col = Vegetation_Zone,
                                  linetype = Vegetation_Zone),
                  linewidth = 0.8) +
        facet_wrap(~Vegetation_Zone) +
        scale_color_manual(values = palcols_zones) +
        theme_bw() +
        labs(x = "Years elapsed",
             y = respn) +
        theme(legend.position = "none")
    
    

    model <- mod_plts
    predicted <- estimate_expectation(model, data = "grid")
    p3 <- plot(predicted, ribbon = list(alpha = 0.2),
               point = list(alpha = 0.5),
               line = list(alpha = 0.9, size = 0.8)) + 
        theme_bw() +
        facet_wrap(~Vegetation_Zone) + 
        scale_color_manual(values = palcols_zones) +
        scale_fill_manual(values = palcols_zones) +
        labs(title = paste("Predicted response:", respn),
             subtitle = "predictions and CIs generated by easystats packages",
             x = "Years elapsed",
             y = respn) +
        theme(legend.position = "none")
    
} else {
    # otherwise, deliver this message:
    cat("### **This statistical model did not run.**")
    cat("\n \nPlease make sure the species or response variable is spelled and capitalized correctly, and that there is adequate data in the file to calculate a trend. \n \nIf that does not help, please copy the error message that appeared above when reaching out to someone else, as it may be useful.")
}
```

### Numerical effects  

```{r, eval = modyes, results = "asis"}

kbl_nmst(mod_anov, 
         caption = paste(respn, " - Type III Analysis of Variance Table, using Kenward-Rogers method"),
         digits = c(4, 4, 1, 2, 2, 4))  

cat("\n \n \n")

cat("Marginal R^2 represents the variance explained by the fixed effects. Conditional R^2 represents the variance explained by the entire model, which includes both fixed and random effects.")

kbl_nmst(rsqds,
         caption = paste(respn, " - Margn'l and Cond'l R^2 values"),
         digits = 3)

cat("\n \n \n")
cat("\n \n \n")

kbl_nmst(mod_trnds, 
         digits = c(4, 4, 4, 2, 4, 4), 
         col.names = c("Veg Zone", "Trend (per year)",
                       "SE", "df", "CI lower", "CI upper"),
         caption = paste(respn, " - Estimated marginal slopes (change per year for each zone)")) 
```


### Graphs  

```{r, warning = FALSE, message = FALSE, eval = modyes}
print(p1)
# print(p2)
print(p3)
```


### Contrasts plot  

If the interaction between vegetation zone and time was significant (p < 0.05), letters are used to represent groups of slopes that are not significantly different (via pairwise comparisons). If the interaction was not significant, this plot represents the marginal slopes but no letters are used.

Some notes from the function. Kenward-Roger method used to estimate denominator degrees of freedom. Confidence level used: 0.95. p-value adjustment: Tukey. alpha = 0.05.  

"If two or more means share the same grouping symbol,
      then we cannot show them to be different.
      But we also did not show them to be the same."

```{r, fig.height = 5, fig.width = 6, eval = modyes}
trnds <- emtrends(mod_plts, 
                  pairwise ~ Vegetation_Zone, 
                  var = "Years_sinceStart")
mod_cld <- cld(trnds, Letters = letters) %>% 
    mutate(.group = str_trim(.group))

mod_cld$Vegetation_Zone <- factor(mod_cld$Vegetation_Zone, levels = zone_factor)

p <- ggplot(mod_cld, aes(x = reorder(Vegetation_Zone, Years_sinceStart.trend),
                         y = Years_sinceStart.trend,
                         label = .group,
                         color = Vegetation_Zone)) +
    geom_hline(yintercept = 0) +
    geom_pointrange(aes(ymin = lower.CL, 
                        ymax = upper.CL),
                    linewidth = 0.8) +
    scale_color_manual(values = palcols_zones) +
    labs(title = paste("Rate of change in", respn, "by zone"),
         x = "Vegetation Zone, low to high rate of change",
         y = "change per year") +
    theme_bw() +
    theme(legend.position = "none")

if(sig_int){
    p <- p +
        geom_text(aes(y = max(mod_cld$upper.CL)),
                  nudge_y = max(mod_cld$upper.CL)/10,
                  color = "black")
}
print(p)
```

Same plot, but x-axis goes from lowest to highest zone.    

```{r, fig.height = 5, fig.width = 6, eval = modyes}
p <- ggplot(mod_cld, aes(x = Vegetation_Zone,
                         y = Years_sinceStart.trend,
                         label = .group,
                         color = Vegetation_Zone)) +
    geom_hline(yintercept = 0) +
    geom_pointrange(aes(ymin = lower.CL, 
                        ymax = upper.CL),
                    linewidth = 0.8) +
    scale_color_manual(values = palcols_zones) +
    labs(title = paste("Rate of change in", respn, "by zone"),
         x = "Vegetation Zone, low to high elevation",
         y = "change per year") +
    theme_bw() +
    theme(legend.position = "none")

if(sig_int){
    p <- p +
        geom_text(aes(y = max(mod_cld$upper.CL)),
                  nudge_y = max(mod_cld$upper.CL)/10,
                  color = "black")
}
print(p)
```



```{r, eval = modyes}
check_singularity(mod_plts)
```


```{r, eval = modyes}
# cleanup before next response var
rm(p, p1, p2, p3,
   mod_plts,
   trends_df,
   rsqds,  
   trnds, mod_cld,
   mod_anov, tidied, sig_int, pdf)
```

