---
title: ''
output: html_document
---

```{r}
library(vegan)
library(ggrepel)
library(ggiraph)
library(plotly)
library(gt)
library(patchwork)
library(metR)
```

## About

These tabs are the multivariate analyses.

```{r mod-multivar-df}
# use data frame where unvegetated 'species' are grouped by nmst categories
# dat_multivar <- dat_nmst_cats

# use the entire data frame - only group up for loading factors of interest
dat_multivar <- dat %>% 
    select(-any_of(c("EMI", "swdiv", "richness")))
```

```{r}
# remove any plots without at least 3 distinct years of data
dat_multivar <- filter(dat_multivar, !(StTrnsPlt %in% not_enough$StTrnsPlt))
```

```{r, results = "asis"}
# if there weren't enough years of data, can't do the rest.
# set up to NOT evaluate code chunks, and spit out a message.
datEnough <- nrow(dat_multivar) > 0

if(!datEnough){
    cat("### **Multivariate analyses were not run.**")
    cat("\n \nThere were not enough rows of data.")
}
```

In this section, we use ordination to visualize and analyze plant communities through plot-level cover data and explanatory variables (e.g., distance to water, elevation, salinity) across sites and marsh zones, when appropriate. Non-metric multidimensional scaling (NMDS) with loading factors (when appropriate) will be used to visualize the data. Loading factors can be key species or groupings (e.g., dominant species, bare, halophytes) or explanatory variables. We also use similarity percentages (SIMPER) to identify which cover categories are contributing to dissimilarity. A PERMANOVA incorporating appropriate explanatory variables (e.g., site, zone) is used to test for differences between the first and last years of monitoring.

### Questions addressed

-   Are there shifts in the vegetation community (as defined by the entire percent-cover matrix) over time? Do these shifts vary by vegetation zone?
    -   addressed with PERMANOVA, comparing first year of monitoring to the 'last' (most recent) year of monitoring.
    -   the first and last year of monitoring at a specific vegetation plot should be considered together. To do this, a single Bray-Curtis dissimilarity was calculated for each vegetation plot and used as the response. Vegetation Zones were then permuted to answer the question of whether community differences between the first and last year were equal across vegetation zones.
    -   if the above test indicated significant differences between Vegetation Zones (p \$ \leq 0.05 \$), the species matrix was split by Vegetation Zone. For each Vegetation Zone, a PERMANOVA was run comparing the community in the first vs. last years. Permutations were restricted so that times (first vs. last) were permuted only within individual plots.
-   Which species/groups contribute most to these shifts?
    -   addressed using SIMPER, comparing first and last year of monitoring data within each zone. SIMPER is only used to follow up on vegetation zones that indicated a significant community difference in the PERMANOVA.
-   Where is the plant community changing, and what characteristics do those areas have in common (e.g. site, zone, distance from water, elevation)?
    -   visualized via NMDS.

### Reserve output choices

The following species/groups were identified by the reserve as important loading factors to display on NMDS outputs.

Sometimes these species do not appear on the plots; usually that is because all data points for that species were 0 or very close to it.

```{r show-userChoice-multivar, results = 'asis'}
choices <- anaSpec_list$multivar$Choice
choices <- choices[!is.na(choices)]
cat("Up to 8 species/groups could be identified specifically in outputs of these multivariate analyses. Reserve choices are: \n")
```

```{r userChoice-2, results = 'asis'}
cat(paste("*", choices), sep="\n")
```

```{r stats-define-matrices, eval = datEnough}
expl <- dat_multivar[, 1:which(names(dat_multivar) == "Total")]
resp <- dat_multivar[, (1 + which(names(dat_multivar) == "Total")):ncol(dat_multivar)]

expl$Zone_Year <- paste(expl$Vegetation_Zone, expl$Year)

env <- expl %>% 
    select(Vegetation_Zone, Year,
           Zone_Year) %>% 
    mutate(across(everything(), as.factor))
```

### Characteristics of multivariate data frame

```{r mod-first-last-years, eval = datEnough}
# find first and last years by zone
start_ends <- dat_multivar %>% 
    group_by(Vegetation_Zone, Year) %>% 
    tally() %>% 
    group_by(Vegetation_Zone) %>% 
    summarize(Start = min(Year),
              End = max(Year))

keeps <- start_ends %>% 
    pivot_longer(Start:End, 
                 names_to = "Time_group",
                 values_to = "Year") 

# subset data frames to first and last years by Zone
expl <- expl %>% 
    left_join(keeps)
fl_index <- !is.na(expl$Time_group)

# make sure we've got enough plots that were sampled in both the start and end years
expl2 <- expl %>% 
    select(StTrnsPlt, Year, Time_group) %>% 
    janitor::tabyl(StTrnsPlt, Time_group)
# sum(expl2$Start) == sum(expl2$End)

expl_pm <- expl[fl_index, ]
resp_pm <- resp[fl_index, ]

expl_pm$Zone_Timegroup <- paste(expl_pm$Vegetation_Zone, expl_pm$Time_group, sep = "; ")

env_pm <- expl_pm %>% 
    select(Vegetation_Zone, Time_group,
           Zone_Timegroup) %>% 
    mutate(across(everything(), as.factor))
```

```{r show-start-end, eval = datEnough}
expl_pm %>% 
    group_by(Vegetation_Zone, Time_group, Year) %>% 
    tally() %>% 
    ungroup() %>% 
    mutate(Year_n = paste0(Year, " (", n, ")")) %>% 
    group_by(Vegetation_Zone, Time_group) %>% 
    summarize(Year_n = paste(unique(Year_n), collapse = " - ")) %>% 
    pivot_wider(names_from = Time_group,
                values_from = Year_n) %>% 
    relocate(Start, .before = End) %>% 
    kbl_nmst(caption = "Years in 'Start' and 'End' time groups for each Vegetation Zone. Number in parentheses is number of samples.",
             align = c("l", "c", "c"))


keeps2 <- expl_pm %>% 
    group_by(Vegetation_Zone, Time_group) %>% 
    summarize(Years = paste(unique(Year), collapse = ", "))
```

```{r show-multivar-df, eval = datEnough}
resp_names <- data.frame(names(resp))
kbl_nmst(resp_names,
         col.names = c(""),
         caption = "Species/Groups included in response matrix") %>%
    scroll_box(width = "300px", height = "400px") 
```

## PERMANOVA

```{r, results = "asis"}
# if there weren't enough years of data, can't do the rest.
# set up to NOT evaluate code chunks, and spit out a message.
datEnough <- nrow(dat_multivar) > 0

if(!datEnough){
    cat("### **Multivariate analyses were not run.**")
    cat("\n \nThere were not enough rows of data.")
}
```

```{r, eval = datEnough}
# response matrix and data frame
resp_mat <- vegdist(resp_pm, method = "bray")
resp_mat.df <- as.data.frame(as.matrix(resp_mat))

# attach names that are plot x time - as both columns and rows
# because it is a symmetric matrix
nms <- paste(expl_pm$StTrnsPlt, expl_pm$Time_group, sep = "_")
names(resp_mat.df) <- nms
resp_mat.df$vegPlot_time <- nms

# pivot and filter out duplicate rows
resp_mat.pivoted <- resp_mat.df %>% 
    pivot_longer(-vegPlot_time,
                 names_to = "vegPlot_time2",
                 values_to = "distance") %>% 
    filter(vegPlot_time != vegPlot_time2)
# split out plots and times, reduce to one row per plot
resp_plotwise <- resp_mat.pivoted %>% 
    separate(vegPlot_time,
             into = c("vegPlot", "time"),
             sep = "_") %>% 
    separate(vegPlot_time2,
             into = c("vegPlot2", "time2"),
             sep = "_") %>% 
    filter(vegPlot == vegPlot2,
           time == "Start",
           time2 == "End") %>% 
    select(vegPlot, distance)

# reattach veg zone
resp_plotwise <- left_join(resp_plotwise,
                           select(stn_tbl, Vegetation_Zone, PlotID_full),
                           by = c("vegPlot" = "PlotID_full")) %>% 
    relocate(Vegetation_Zone)
```

```{r, eval = datEnough}
lm_zonedist <- lm(distance ~ Vegetation_Zone, data = resp_plotwise)

# overall F for model (same as F for Vegetation_Zone, since it's one-way - really this is our interaction, if we would have kept start/end separate) 
# are there differences between zones in their response to time?
observed_F <- summary(lm_zonedist)$fstatistic[["value"]]
```

```{r, eval = datEnough}
nperm <- 999
permuted_F <- numeric(nperm)

set.seed(1223)
shuffledSet <- shuffleSet(nrow(resp_plotwise), nperm)

for(i in seq_len(nperm)){
    shuffled <- resp_plotwise %>% 
        mutate(Vegetation_Zone = resp_plotwise$Vegetation_Zone[shuffledSet[i,]])
    
    lm_shuffled <- lm(distance ~ Vegetation_Zone, data = shuffled)

    # overall F
    permuted_F[i] <- summary(lm_shuffled)$fstatistic[[1]]
}
```

```{r, results = 'asis', eval = datEnough}
permuted_F <- c(observed_F, permuted_F)

pval_F <- sum(permuted_F >= observed_F) / length(permuted_F)

cat("\n\nH0: community distance between time points is equal across all vegetation zones",
    "\n\nHa: at least one vegetation zone has different community change than at least one other zone"
)

runZoneByZone <- pval_F <= 0.1 & datEnough == TRUE
runOverall <- pval_F > 0.1 & datEnough == TRUE
```


Zone-wise PERMANOVAs are run below if the p-value from the test above is <= 0.10.  

```{r, eval = datEnough, fig.width = 7, fig.height = 3}
permtab <- data.frame(a = c("Number of permutations",
                            "Observed F statistic",
                            "Permutation-based p-value"),
                      b = c(nperm, round(observed_F, 3), round(pval_F, 4)))

permtabgt <- gt(permtab) %>% 
    cols_label(
        a = "PERMANOVA output",
        b = ""
    ) %>% 
    fmt_number(rows = 1, decimals = 0) %>% 
    fmt_number(rows = c(2:3), decimals = 3)

permutedFdf <- data.frame(rep = 1:length(permuted_F),
                          permF = permuted_F)

permplot <- ggplot(permutedFdf) +
    geom_histogram(aes(x = permuted_F),
                   fill = "gray80",
                   col = "gray30",
                   bins = 30) +
    geom_vline(xintercept = observed_F, 
               col = "red", 
               linetype = "dashed", 
               size = 1) +
    theme_bw() +
    labs(title = "Permutation distribution",
         subtitle = "red line = observed F value",
         x = "F")

# patchwork to display together
wrap_table(permtabgt) + plot_spacer() + permplot +
    plot_layout(widths = c(3, 1, 4))
```

#### Check for homogeneity of variance  

Because the permutation test above was performed on a univariate response (Bray-Curtis distance between 'start' and 'end' for each plot), a Levene's test is used here to check for homogeneity of variances between groups. If p < 0.05, this assumption is violated and results of the permutational test above should be interpreted with caution. A violin plot to visualize distances by vegetation zone is also below, whether or not the Levene's test is significant.    

```{r, eval = datEnough, fig.width = 5, fig.height = 4}
levene_out <- car::leveneTest(distance ~ Vegetation_Zone, data = resp_plotwise)
kbl_nmst(levene_out,
         digits = c(0, 2, 4),
         caption = "Levene's test for homogeneity of variances; center = median")


p_violin <- ggplot(resp_plotwise,
       aes(x = Vegetation_Zone,
           y = distance,
           fill = Vegetation_Zone)) +
    geom_violin(draw_quantiles = c(0.25, 0.5, 0.75),
                alpha = 0.7) +
    scale_fill_manual(values = palcols_zones) +
    theme_bw() +
    theme(legend.position = "none") +
    theme(axis.text.x = element_text(angle = 45,
                                     hjust = 1)) +
    labs(title = "Distribution of distance by vegetation zone",
         subtitle = "horizontal lines represent 25th percentile, median, and 75th percentile for each group",
         x = NULL,
         y = "Bray-Curtis distance")

p_violin
```



### Zone-wise

If the above permutational test for differences in community response between zones was significant or close to it (p <= 0.10), zone-wise tests are below. Otherwise, all zones were combined for one test of start vs. end community change.

Plots are paired with each other - i.e. time points can only be swapped within a plot.  

As above, a test for homogeneity of group variances is conducted, this time for start vs. end within each vegetation zone, controlling for plot. Because the response in this instance is multivariate, the `vegan::betadisper()` function, the R analogue to the PERMDISP test, is used. PERMDISP is the multivariate analogue of Levene's test for homogeneity of variances.  

```{r, results = 'asis'}
set.seed(1223)

# smaller data frame
expl_pm2 <- expl_pm %>% 
    select(Vegetation_Zone, StTrnsPlt, Time_group)
# set up veg zones
veg_zones <- levels(expl_pm2$Vegetation_Zone)

pmvs <- list()
simps <- list()
betadispers <- list()
dispers_toplo <- list()  # to make dispersion plots  
```

```{r, results = 'asis', eval = runZoneByZone}

for(i in seq_along(veg_zones)){
    indices <- which(expl_pm2$Vegetation_Zone == veg_zones[i])
    
    resp_zn <- resp_pm[indices, ]
    expl_zn <- expl_pm2[indices, ]
    
    pmvt <- adonis2(resp_zn ~ Time_group, data = expl_zn,
                    control = how(blocks = expl_zn$StTrnsPlt), # don't move plots, just swap time stamps within where they are
                    method = "bray",
                    by = "margin")
    
    # PERMDISP
    disper.obj <- betadisper(vegdist(resp_zn, method = "bray"),
                             group = expl_zn$Time_group)
    disper.test <- permutest(disper.obj, control = how(blocks = expl_zn$StTrnsPlt))
    
    disper.test2 <- disper.test$tab
    
    
    disper.df <- data.frame(
        Start.dispersion = disper.obj$group.distances["Start"],
        End.dispersion = disper.obj$group.distances["End"],
        NumDf = disper.test2[1, "Df"],
        DenDf = disper.test2[2, "Df"],
        F.val = disper.test2[1, "F"],
        N.Perm = disper.test2[1, "N.Perm"],
        p.val = disper.test2[1, "Pr(>F)"])
    
    betadispers[[i]] <- disper.df
    names(betadispers)[[i]] <- as.character(veg_zones[i])
    
    dispers_toplo[[i]] <- disper.obj
    names(dispers_toplo)[[i]] <- as.character(veg_zones[i])
    
    # save formatted output to list
    pmvs[[i]] <- broom::tidy(pmvt)
    names(pmvs)[[i]] <- as.character(veg_zones[i])
    
    # if permanova p-value exists and was below 0.2, run SIMPER
    if (length(pmvt$`Pr(>F)`[row.names(pmvt) == "Time_group"]) == 1) {
        
        # if p-value is less than 0.2, create simper
        if (pmvt$`Pr(>F)`[row.names(pmvt) == "Time_group"] <= 0.200) {
            
            # run simper
            simp <- simper(resp_zn, expl_zn$Time_group)
            
            # format output
            simp_df <- summary(simp, ordered = TRUE)$Start_End[1:6, c("average", "sd", "cumsum", "p", "ava", "avb")]
            # in case there are fewer than 6 species, get rid of empty rows
            simp_df <- janitor::remove_empty(simp_df, "rows")
            simp_smaller <- data.frame(lapply(simp_df, round, 4))
            simp_smaller$ava <- round(simp_smaller$ava, 1)
            simp_smaller$avb <- round(simp_smaller$avb, 1)
            row.names(simp_smaller) <- row.names(simp_df)
            names(simp_smaller) <- c("average", "sd", "cumulative", "p", "mean_start", "mean_end")
            
            # save output to list
            simps[[i]] <- simp_smaller
            
            # if permanova wasn't significant, save a placeholder in the list   
        } else {
            simps[[i]] <- data.frame("None" = "SIMPER not run")
        }
        
        # if p-value doesn't exist, save a placeholder in the list
    } else {
        simps[[i]] <- data.frame("None" = "SIMPER not run")
    }
    
    names(simps)[[i]] <- as.character(veg_zones[i])
}
```

```{r, results = 'asis', eval = runZoneByZone}

# PERMANOVA outputs

pmvs_summ <- bind_rows(pmvs, .id = "Vegetation Zone") %>% 
    filter(term == "Time_group") %>% 
    select('Vegetation Zone', R2, p.value) %>% 
    mutate(p.value_Bonferonni = p.adjust(p.value, method = "bonferroni"))

print(kbl_nmst(pmvs_summ,
               digits = 4,
               caption = "Summary of pairwise PERMANOVAs testing community difference between first and most recent years of monitoring"))


# PERMDISP outputs
betadispers_summ <- bind_rows(betadispers, .id = "Vegetation Zone")
rownames(betadispers_summ) <- NULL

print(kbl_nmst(betadispers_summ,
      digits = c(0, 3, 3, 0, 0, 1, 0, 4),
      caption = "Summary of PERMDISP outputs for each zone"))


```

<details><summary>Click to view plots of dispersion for each zone</summary>

```{r, results = 'asis', eval = runZoneByZone, fig.width = 4, fig.height = 3}
for(i in seq_along(dispers_toplo)){
    plot(dispers_toplo[[i]],
         main = names(dispers_toplo)[[i]])
}
```

</details>  

---



```{r, results = "asis", eval = runOverall}
cat("\n\n")

cat("\n\nNO DIFFERENCE BETWEEN VEGETATION ZONES\n")
    
cat("\nCONDUCTING START vs. END ANALYSIS ACROSS ALL ZONES\n-------------------\n\n")

pmv <- adonis2(resp_pm ~ Time_group, data = expl_pm2,
               control = how(blocks = expl_pm2$StTrnsPlt), # don't move plots, just swap time stamps within where the plot data rows are
               method = "bray",
               by = "margin")

# save output to proper list for printing later
pmvs_summ <- broom::tidy(pmv)

# PERMDISP
disper.obj <- betadisper(vegdist(resp_pm, method = "bray"),
                         group = expl_pm2$Time_group)
disper.test <- permutest(disper.obj, control = how(blocks = expl_pm2$StTrnsPlt))
disper.test2 <- disper.test$tab
disper.df <- data.frame(
    Start.dispersion = disper.obj$group.distances["Start"],
    End.dispersion = disper.obj$group.distances["End"],
    NumDf = disper.test2[1, "Df"],
    DenDf = disper.test2[2, "Df"],
    F.val = disper.test2[1, "F"],
    N.Perm = disper.test2[1, "N.Perm"],
    p.val = disper.test2[1, "Pr(>F)"])
rownames(disper.df) <- NULL

print(kbl_nmst(pmv,
               digits = c(0, 2, 4, 2, 4),
               caption = "PERMANOVA table, community difference at start vs. end across all zones") )

print(kbl_nmst(disper.df,
      digits = c(3, 3, 0, 0, 1, 0, 4),
      caption = "PERMDISP output, start vs. end across all zones"))


# run a simper if this PERMANOVA had p <= 0.02
if(pmv$`Pr(>F)`[row.names(pmv) == "Time_group"] <= 0.200){
    # run simper
    simp <- simper(resp_pm, expl_pm$Time_group)
    
    # format output
    simp_df <- summary(simp, ordered = TRUE)$Start_End[1:6, c("average", "sd", "cumsum", "p", "ava", "avb")]
    # in case there are fewer than 6 species, get rid of empty rows
    simp_df <- janitor::remove_empty(simp_df, "rows")
    simp_smaller <- data.frame(lapply(simp_df, round, 4))
    simp_smaller$ava <- round(simp_smaller$ava, 1)
    simp_smaller$avb <- round(simp_smaller$avb, 1)
    row.names(simp_smaller) <- row.names(simp_df)
    names(simp_smaller) <- c("average", "sd", "cumulative", "p", "mean_start", "mean_end")
    
    # save output to list
    simps[[1]] <- simp_smaller
} else {
    simps[[1]] <- data.frame("None" = "SIMPER not run")
}

names(simps)[[1]] <- "Across all zones"
```

## SIMPER

```{r, results = "asis"}
# if there weren't enough years of data, can't do the rest.
# set up to NOT evaluate code chunks, and spit out a message.
datEnough <- nrow(dat_multivar) > 0

if(!datEnough){
    cat("### **Multivariate analyses were not run.**")
    cat("\n \nThere were not enough rows of data.")
}
```

If the PERMANOVA for a single zone had p-value \< 0.2; *not* adjusted for multiple comparisons due to exploratory nature of these analyses, and note this does **not** represent a significant difference, a SIMPER was performed. **Please check the 'Zone-wise' table in the PERMANOVA tab for the p-value, and be cautious in interpretation.**

The top 6 species in output are below.

"average" is the average contribution of that species to the Bray-Curtis distance between the two groups (note, this is *not* expressed in % and the column does not total to 1); "sd" is the standard deviation of the species' contribution. "cumulative" is the cumulative % contribution for this species and all those above it in the table. Typically people only report species up to the one that brings "cumulative" over 0.7. "p" is a p-value for that species based on permutation tests. "mean_start" is the mean cover of that species in the starting year(s), and "mean_end" is the mean cover of the species in the last year(s) of monitoring.

```{r show-simper, results = 'asis', eval = datEnough}
for(i in seq_along(simps)){
    print(kbl_nmst(x = simps[[i]], 
                   caption = paste0(names(simps)[i], " SIMPER results; PERMANOVA p = ", round(pmvs_summ$p.value[i], 3)),
                   digits = 4)
    )
}
```

## NMDS

```{r, results = "asis"}
# if there weren't enough years of data, can't do the rest.
# set up to NOT evaluate code chunks, and spit out a message.
datEnough <- nrow(dat_multivar) > 0

if(!datEnough){
    cat("### **Multivariate analyses were not run.**")
    cat("\n \nThere were not enough rows of data.")
}
```

```{r stats-run-nmds, eval = datEnough}
set.seed(1226)    
mds_resp <- metaMDS(resp_pm, distance = "bray", k = 3,
                        autotransform = FALSE, trace = 0)

    data.scores <- as.data.frame(scores(mds_resp)$sites)
    species.scores <- as.data.frame(scores(mds_resp)$species)
    species.scores$species <- rownames(species.scores) 
    
    # only keep the species/groups listed as reserve choices, to reduce clutter on graphic
    # first get everything to group
    species.scores <- species.scores %>% 
        left_join(species_info, by = c("species" = "Species")) %>% 
        mutate(Species_or_Group = case_when(species %in% choices ~ species,
                                            Plant_Categories %in% choices ~ Plant_Categories,
                                            NMST_Groupings %in% choices ~ NMST_Groupings,
                                            Cover_Categories %in% choices ~ Cover_Categories,
                                            .default = "Other")) %>% 
        filter(Species_or_Group %in% choices) %>% 
        select(species = Species_or_Group,
               NMDS1,
               NMDS2,
               NMDS3)
    
    # then find centroids for the species-or-groups
    species.scores <- aggregate(cbind(NMDS1, NMDS2, NMDS3) ~ species, data = species.scores, FUN = mean)
    
    # data.scores$Vegetation_Zone <- expl$Vegetation_Zone
    # data.scores$Year <- expl$Year
    data.scores <- data.scores %>%
        mutate(Zone = expl_pm$Vegetation_Zone,
               Year = as.factor(expl_pm$Year),
               Time_group = as.factor(expl_pm$Time_group),
               Zone_Timegroup = as.factor(expl_pm$Zone_Timegroup)) %>% 
        separate(Zone, into = c("Zone_abbrev", "Zone_full"),
                 sep = "-")  
```

------------------------------------------------------------------------

Final 3-dimensional NMDS stress was **`r ifelse(datEnough, round(mds_resp$stress, 4), "NMDS not run")`**.

Rules of thumb for interpreting stress, based on the sources below, are:

```{r}

stress_interp <- tribble(
    ~"Stress", ~"Quality of ordination",
    "< 0.05" , "Excellent" ,
"0.05 ‒ 0.10" , "Good; no need to increase number of dimensions" ,
"0.10 ‒ 0.20" , "Useable, but do not place too much reliance on details, as a higher-dimensional ordination could be different" ,  
"0.20 ‒ 0.30" , "Interpretation problems start, and especially in the upper range of the interval could be dangerously misleading" ,
"> 0.30" , "Poor presentation; essentially random placement of points. Consider increasing number of dimensions"
)

DT::datatable(stress_interp,
              rownames = FALSE,
              width = "600px",
              options = list(dom = "t",
                             columnDefs = list(
                                 list(className = 'dt-right', targets = 0),
                                 list(width = '25%', targets = 0)
                             )))
```

**For more information on NMDS:**

Clarke, K. R. (1993). Non-parametric multivariate analyses of changes in community structure. Australian Journal of Ecology, 18(1), Article 1. <https://doi.org/10.1111/j.1442-9993.1993.tb00438.x>

Clarke, K. R., & Warwick, R. M. (2001). Change in Marine Communities: An Approach to Statistical Analysis and Interpretation, 2nd ed. - Chapter 5 focuses on NMDS.

Zuur, A. F., Ieno, E. N., & Smith, G. M. (2007). Analysing ecological data. Springer. - Chapter 15 for NMDS.

#### 3-d NMDS plot

```{r plot-nmds-3d, eval = datEnough, fig.width = 6, fig.height = 6}
# set up zone colors
zone_cols <- as.vector(palcols_zones)
names(zone_cols) <- NULL

# main plot - points for each zone
# (no centroids)
fig <- plot_ly() %>% 
    add_trace(data = data.scores,
              type = "scatter3d",
              mode = "markers",
              x = ~NMDS1,
              y = ~NMDS2,
              z = ~NMDS3,
              color = ~Zone_abbrev,
              colors = zone_cols,
              symbol = ~Time_group,
              symbols = c("circle", "circle-open"),
              marker = list(size = 5,
                            opacity = 0.7)
    ) 

# now add lines, text, and cross-hatches for each identified loading
for (i in 1:nrow(species.scores)) {
    fig <- fig %>% 
        add_paths(
            x = c(0, species.scores$NMDS1[i]),
            y = c(0, species.scores$NMDS2[i]),
            z = c(0, species.scores$NMDS3[i]),
            # type = "scatter3d",
            mode = "lines",
            line = list(color = "red", width = 4),
            showlegend = FALSE
        )  |>
        add_text(
            x = species.scores$NMDS1[i],
            y = species.scores$NMDS2[i],
            z = species.scores$NMDS3[i],
            text = species.scores$species[i],
            textfont = list(color = "red", size = 12),
            showlegend = FALSE
        )  %>% 
        add_trace(
            type = "scatter3d",
            mode = "markers",
            x = species.scores$NMDS1[i],
            y = species.scores$NMDS2[i],
            z = species.scores$NMDS3[i],
            marker = list(size = 6,
                          color = "red",
                          symbol = "cross"),
            showlegend = FALSE
        )
        
}

# save the plot to another file
# NMDS_outname <- paste0("NMDS-3d_", params$file_code, "_", Sys.Date(), ".html")
# saveWidget(fig, file = here::here("output", NMDS_outname), selfcontained = TRUE)

# print the plot
fig
```

#### 2-dimensional NMDS plot

(first two axes only)

#### How to read these NMDS plots

-   Large points represent the centroid for each Vegetation Zone/Time period combination. They are labelled with the Vegetation Zone abbreviation and 'Start' or 'End'. Additionally, they are colored by Vegetation Zone and shaped by 'Start' vs. 'End' time periods. Upon hover, the full Vegetation Zone name and year(s) represented will be provided.\
-   Black arrows and text represent Species or Categories of vegetation (e.g. H-Halophyte) identified by the reserve. Hovering over the arrow will provide the Species or Category name.\
-   Small points represent individual plots and are colored by Vegetation Zone and shaped by 'Start' vs. 'End' time periods (shapes and colors match the large points above); hovering does not yield output for points.\
-   Some identified loading factors may not appear as arrows; this is usually because all data points for that species were 0. They may also have been too low to have produced species scores in the NMDS.

```{r stats-prep-nmds-groups, eval = datEnough}
# get centroids for each time/zone group
# https://stackoverflow.com/a/47523181
cent <- aggregate(cbind(NMDS1, NMDS2, NMDS3) ~ Zone_Timegroup, data = data.scores, FUN = mean)

en_coord_cat <- cent %>%
    separate(Zone_Timegroup, into = c("Zone", "Time_group"),
             sep = "; ") %>%
    left_join(keeps2,
              by = c("Zone" = "Vegetation_Zone",
                     "Time_group")) %>% 
    separate(Zone, into = c("Zone_abbrev", "Zone_full"),
             sep = "-") %>%
    mutate(Zone_Timegroup = paste(Zone_abbrev, Time_group),
           Zone_Years = paste(Zone_full, Years, sep = ": "))
```

```{r stats-graph-nmds-centroids, fig.width = 6, fig.height = 5, eval = datEnough}
p <- plot_nmds() 
girafe(ggobj = p)
```

### Contour plots  

The below plots, rather than using arrows for individual species or vegetation groups, create contours for the specific values of each. Contours are labeled with % cover values and are fit as a spline-based surface using `vegan::ordisurf()`.      

<details><summary>Click to expand contour plots.</summary>

```{r, fig.width = 6, fig.height = 6}
spps <- paste0(species.scores$species)
for(i in seq_along(spps)){
    # Construct formula dynamically with backticks
  formula <- as.formula(paste0("mds_resp ~ `", spps[i], "`"))
  
  # Pass the formula to ordisurf
  ordi <- ordisurf(formula, resp_pm, plot = FALSE)
  
  grid.obj <- with(ordi$grid, expand.grid(x = x, y = y))
  grid.obj$z <- as.vector(ordi$grid$z)
  p_cont <- plot_nmds_contours() +
      labs(subtitle = paste0("Contours for: ", spps[i], "\nLarge points represent Zone/Time centroids"))
  print(p_cont)

  # plot(ordi, main = spps[i], xlab = "NMDS1", ylab = "NMDS2")
}
```

</details>
