---
title: ''
output: html_document
---

Modifying multivariate analyses to only include first and last year.  

```{r}
library(tidyverse)
library(vegan)

```




```{r}
# test data
testdf <- data.frame("veg_zone" = rep(c("Low Marsh", "High Marsh"), each = 4),
                     "veg_plot" = rep(c("L.1", "L.2", "H.1", "H.2"), each = 2),
                     "time" = rep(c("time 1", "time 2"), times = 4))
```

Move time within veg plot, and move plots around. Unfortunately this keeps veg_zone associated with each veg_plot.  

```{r}
# option 1 which I think doesn't do what I want
ctrl <- how(
    plots = Plots(strata = testdf$veg_plot, type = "free"),   # permute veg plot samples as a unit (both time groups for a plot together get permuted to the same zone)
    within = Within(type = "free"),  # samples (time group) can be permuted within veg plots
    )

set.seed(1222)
testdf[shuffle(testdf, control = ctrl), ]
# permuted plots and time, but veg zone went with it
```

Shuffles time stamp within veg plots only. This is okay when we're looking at a single veg_zone, so it doesn't matter that zone isn't getting shuffled. It keeps the community data with each veg_plot and only shuffles time stamps within veg_plot.  

```{r}
ctrl2 <- how(blocks = testdf$veg_plot)

testdf[shuffle(testdf, control = ctrl2), ]
```


In an ideal world, we would permute both the time stamp and the veg_zone with respect to veg plot. I couldn't figure out how to do this without sequentially manually permuting the data, and then I'd have to figure out how to calculate the F statistic. And .... how does that work???  

```{r}
# time stamps within plot
ctrl_step1 <- how(blocks = testdf$veg_plot)

# plots between zones
ctrl_step2 <- how(
    plots = Plots(strata = testdf$veg_plot, type = "free"),   # permute veg plot samples as a unit (both time groups for a plot together get permuted to the same zone)
    within = Within(type = "none"),  # samples (time group) cannot be permuted
    )

set.seed(2025)
intermed <- testdf[shuffle(testdf, control = ctrl_step1), ]
intermed[shuffle(intermed, control = ctrl_step2), ]
```

The dissimilarity betweent time points for a plot will always be the same if we only have a two time points to compare. The only difference is what zone we've labelled that plot with. 

So if I don't worry about swapping time points, and only swap zones around, that should give me the same answer regarding the interaction as calculating the distance for each plot and running it in a univariate way did. This will ONLY tell me about the interaction, and NOT about the main effect of time point, if we're not permuting time point.  

```{r}
testdf[shuffle(testdf, control = ctrl_step2), ]
```



# one more attempt  

holy cow this is what I wanted in the first place  

```{r}
ctrl3 <- how(
    plots = Plots(strata = testdf$veg_plot, type = "free"),   # permute veg plot samples as a unit (both time groups for a plot together get permuted to the same zone; thish brings veg_zone along with the plot)
    within = Within(type = "free"),  # samples (time group) can be permuted
    )

testdf[shuffle(testdf, control = ctrl3),]
```

# test some things  

Using GND data - running from the beginning of 99_Veg_analyses, through data frame prep for the permanova in 004 Multivariate, via those scripts. Then start here.  

Using resp_pm and expl_pm - these are the datasets subsetted to start and finish.  

Turns out you need to arrange the data frame properly first (or at least that helps verify that the permutations are occurring correctly - maybe it was swapping correctly before but I didn't realize because the plots' different time points were not next to each other in the data frame).  And I need to do that before splitting it into explanatory and response. I have taken care of this now in the multivariate data frame.   

```{r}
ctrl_pm <- how(
    plots = Plots(strata = expl_pm$StTrnsPlt, type = "free"),   # permute veg plot samples as a unit (both time groups for a plot together get permuted to the same zone; thish brings veg_zone along with the plot)
    within = Within(type = "free"),  # samples (time group) can be permuted
    nperm = 999
    )

head(expl_pm[shuffle(expl_pm, control = ctrl_pm), c("Vegetation_Zone", "StTrnsPlt", "Time_group")], 10)

```

## permanova 

```{r}
pmvt <- adonis2(resp_pm ~ Vegetation_Zone*Time_group, data = expl_pm,
                    control = ctrl_pm,
                    method = "bray",
                    by = "margin")

# veg zone main effect
pmvt2 <- adonis2(resp_pm ~ Vegetation_Zone*Time_group, data = expl_pm,
                    control = ctrl_pm,
                    method = "bray",
                    by = "terms")

pmvt3 <- adonis2(resp_pm ~ Time_group*Vegetation_Zone, data = expl_pm,
                    control = ctrl_pm,
                    method = "bray",
                    by = "terms")

pmvt4 <- adonis2(resp_pm ~ Vegetation_Zone*Time_group, data = expl_pm,
                    control = ctrl_pm,
                    method = "bray")
```



# ACTUALLY THIS -  

Calculate distance for each plot b/t start and end. then it's univariate. Still permute; is valid to permute raw data, and even recommended when it's a one-way situation (https://learninghub.primer-e.com/books/permanova-for-primer-guide-to-software-and-statistical-methods/page/117-methods-of-permutations)  


## 1. Calculate distances for each plot  

Calculate the whole darn distance matrix and then weed it down to plot-wise comparisons.  

```{r}
# response matrix and data frame
resp_mat <- vegdist(resp_pm, method = "bray")
resp_mat.df <- as.data.frame(as.matrix(resp_mat))

# attach names that are plot x time - as both columns and rows
# because it is a symmetric matrix
nms <- paste(expl_pm$StTrnsPlt, expl_pm$Time_group, sep = "_")
names(resp_mat.df) <- nms
resp_mat.df$vegPlot_time <- nms

# pivot and filter out duplicate rows
resp_mat.pivoted <- resp_mat.df %>% 
    pivot_longer(-vegPlot_time,
                 names_to = "vegPlot_time2",
                 values_to = "distance") %>% 
    filter(vegPlot_time != vegPlot_time2)
# split out plots and times, reduct to one row per plot
resp_plotwise <- resp_mat.pivoted %>% 
    separate(vegPlot_time,
             into = c("vegPlot", "time"),
             sep = "_") %>% 
    separate(vegPlot_time2,
             into = c("vegPlot2", "time2"),
             sep = "_") %>% 
    filter(vegPlot == vegPlot2,
           time == "Start",
           time2 == "End") %>% 
    select(vegPlot, distance)

# reattach veg zone
resp_plotwise <- left_join(resp_plotwise,
                           select(stn_tbl, Vegetation_Zone, PlotID_full),
                           by = c("vegPlot" = "PlotID_full")) %>% 
    relocate(Vegetation_Zone)
```

So now the data frame has one row per vegetation plot, and 3 columns: Vegetation Zone, Plot ID, and Bray-Curtis distance between first and last years.  

## Calculate observed F statistic  

```{r}
lm_zonedist <- lm(distance ~ Vegetation_Zone, data = resp_plotwise)

# overall F for model (same as F for Vegetation_Zone, since it's one-way - really this is our interaction, if we would have kept start/end separate) 
# are there differences between zones in their response to time?
observed_F <- summary(lm_zonedist)$fstatistic[["value"]]
```


## Permute vegetation zone and get permutation distribution  

```{r}
nperm <- 999
permuted_F <- numeric(nperm)

set.seed(1223)
shuffledSet <- shuffleSet(nrow(resp_plotwise), nperm)

for(i in seq_len(nperm)){
    shuffled <- resp_plotwise %>% 
        mutate(Vegetation_Zone = resp_plotwise$Vegetation_Zone[shuffledSet[i,]])
    
    lm_shuffled <- lm(distance ~ Vegetation_Zone, data = shuffled)

    # overall F
    permuted_F[i] <- summary(lm_shuffled)$fstatistic[[1]]
}
```

## Combine, visualize, calculate p-vals  

```{r}
permuted_F <- c(observed_F, permuted_F)

pval_F <- sum(permuted_F >= observed_F) / length(permuted_F)

hist(permuted_F, breaks = 30,
       main = "Permutation distribution")
abline(v = observed_F, col = "red", lwd = 2, lty = 2)

cat("Observed F: ", round(observed_F, 3))
cat("Permutation-based p-value: ", round(pval_F, 4))

cat("H0: community distance between time points is equal across all vegetation zones",
    "\nHa: at least one vegetation zone has different community change than at least one other zone",
    "\nNumber of Permutations: ", nperm, 
  "\n\nObserved F: ", round(observed_F, 3), "\nPermutation-based p-value: ", round(pval_F, 4), "\n"
)

```


# Follow-up  

If the test above was significant, carry out zone-wise tests. If not, check for overall start vs. end community change.  

Plots are paired with each other - i.e. time points can only be swapped within a plot.  

```{r}
set.seed(1223)

# smaller data frame
expl_pm2 <- expl_pm %>% 
    select(Vegetation_Zone, StTrnsPlt, Time_group)
# set up veg zones
veg_zones <- levels(expl_pm2$Vegetation_Zone)

if(pval_F <= 0.05){
    
    cat("\n\nZONE-BY-ZONE ANALYSIS\n-------------------\n\n")
    
    for(i in seq_along(veg_zones)){
        indices <- which(expl_pm2$Vegetation_Zone == veg_zones[i])
        
        resp_zn <- resp_pm[indices, ]
        expl_zn <- expl_pm2[indices, ]
        
        cat("Zone: ", veg_zones[i], "\n\n")
        print(adonis2(resp_zn ~ Time_group, data = expl_zn,
                      control = how(blocks = expl_zn$StTrnsPlt), # don't move plots, just swap time stamps within where they are
                      method = "bray",
                      by = "margin"))
        
        cat("\n--------------------\n\n")
    }
    
} else {
    
    cat("\n\nNO DIFFERENCE BETWEEN VEGETATION ZONES\nCONDUCTING START vs. END ANALYSIS ACROSS ALL ZONES\n-------------------\n\n")
    
    print(adonis2(resp_pm ~ Time_group, data = expl_pm2,
                      control = how(blocks = expl_pm2$StTrnsPlt), # don't move plots, just swap time stamps within where they are
                      method = "bray",
                      by = "margin"))
    
    
}
```





# THIS IS WHAT I NEED  

Well, in terms of permutations it is, but truly, permuting time within plot - since there are only 2 time points - won't be any different than just calculating a distance between time points per plot and doing a one-way analysis.  

```{r}
# two steps:
# first, permute vegetation zones for plots
# by only using one row per plot
testdf2 <- testdf %>% 
    select(veg_zone, veg_plot) %>% 
    distinct()
permuted_zones <- testdf2$veg_zone[shuffle(testdf2$veg_zone)]
permuted_zonesdf <- data.frame("permuted_zone" = permuted_zones,
                               "veg_plot" = testdf2$veg_plot)


# permute time stamps within plot 
testdf3 <- testdf %>% 
    select(veg_plot, time)
ctrl_times <- how(within = Within(type = "free"),
                  plots = Plots(strata = testdf3$veg_plot,
                                type = "none"))
permuted_timesdf <- testdf3[shuffle(testdf3, control = ctrl_times), ]



# join permuted plot timestamps to permuted zones
permuted_all <- left_join(permuted_timesdf, permuted_zonesdf,
                          by = "veg_plot")
permuted_all
```


Set a seed, then keep running the chunk above, to see how things come out. It seems to work.  

Next thing then is to figure out sums of squares and degrees of freedom - since this isn't a standard design in permute, vegan won't figure it out for me.  


What if I just assign all plots to the same block? will it let me shuffle plots around relative to vegetation zones?  

```{r}
testdf4 <- testdf %>% 
    mutate(block = 1)
ctrl_blocks <- how(blocks = testdf4$block,
                   plots = Plots(strata = testdf4$veg_plot,
                                 type = "free"),
                   within = Within(type = "free"))

testdf4[shuffle(testdf4, control = ctrl_blocks), ]
```


# nope  

```{r}
ctrl <- how(
    plots = Plots(strata = expl_pm$StTrnsPlt, type = "free"),   # permute veg plot samples as a unit (both time groups for a plot together get permuted to the same zone)
    within = Within(type = "free"),  # samples (time group) can be permuted within veg plots
    )

# see what that setup looks like
ctrl

```

```{r}
# make a backup of the data and then do some column selection for example permutations
expl_pm_backup <- expl_pm
expl_pm <- expl_pm %>% 
    select(Vegetation_Zone, StTrnsPlt,
           Year, Time_group)


# view a few permutations
set.seed(1220)
permutations <- shuffleSet(n = nrow(expl_pm), control = ctrl, nset = 5)

# Inspect the original data and permutations
original_data <- expl_pm[1:10, ]  # First 10 rows of the dataset for demonstration
perm_examples <- lapply(1:nrow(permutations), function(i) {
  expl_pm[permutations[i, ], ]
})

# Print original data and two permuted examples
cat("Original Data (first 10 rows):\n")
expl_pm %>% 
    mutate(rownum = row_number()) %>% 
    filter(str_starts(StTrnsPlt, "spalt")) %>% 
    arrange(StTrnsPlt) %>% 
    head(10)

cat("\nPermuted Example 1:\n")
perm_examples[[1]] %>% 
    as.data.frame() %>% 
    mutate(rownum = row_number()) %>% 
    filter(str_starts(StTrnsPlt, "spalt")) %>% 
    arrange(StTrnsPlt) %>% 
    head(10)


cat("\nPermuted Example 2:\n")
perm_examples[[2]] %>% 
    as.data.frame() %>% 
    mutate(rownum = row_number()) %>% 
    filter(str_starts(StTrnsPlt, "spalt")) %>% 
    arrange(StTrnsPlt) %>% 
    head(10)


# replace expl_pm with its original
expl_pm <- expl_pm_backup
```


```{r stats-run-permanova, eval = datEnough}
pmv <- adonis2(resp_pm ~ Vegetation_Zone*Time_group, data = expl_pm,
               by = "terms",
        method = "bray",
        permutations = 5)

attributes(pmv)$F.perm

testpmv <- adonis2(resp_pm ~ Vegetation_Zone + Time_group, 
                   data = expl_pm, by = "terms", method = "bray")

kbl_nmst(pmv,
      digits = c(0, 2, 4, 2, 4),
      caption = "PERMANOVA table") 
```


```{r stats-perm-posthoc, eval = datEnough}
# for both follow-ups, want each zone on its own, start vs. end

zones <- unique(expl_pm$Vegetation_Zone)
pmvs <- list()
simps <- list()

for(i in seq_along(zones)){
    # subset data to this zone
    zn = zones[i]
    tmp_index <- which(expl_pm$Vegetation_Zone == zn)
    tmp_resp <- resp_pm[tmp_index, ]
    tmp_expl <- expl_pm[tmp_index, ]
    
    # run permanova
    pmvt <- adonis2(tmp_resp ~ Time_group, data = tmp_expl,
                    method = "bray", by = "terms")
    
    # save formatted output to list
    pmvs[[i]] <- broom::tidy(pmvt)
    names(pmvs)[[i]] <- as.character(zn)
    
    # if permanova p-value was below 0.2, run SIMPER
    if(pmvt$`Pr(>F)`[row.names(pmvt) == "Time_group"] < 0.200){
        # run simper
        simp <- simper(tmp_resp, tmp_expl$Time_group)
        
        # format output
        simp_df <- summary(simp, ordered = TRUE)$Start_End[1:6, c("average", "sd", "cumsum", "p", "ava", "avb")]
        # in case there are fewer than 6 species, get rid of empty rows
        simp_df <- janitor::remove_empty(simp_df, "rows")
        simp_smaller <- data.frame(lapply(simp_df, round, 4))
        simp_smaller$ava <- round(simp_smaller$ava, 1)
        simp_smaller$avb <- round(simp_smaller$avb, 1)
        row.names(simp_smaller) <- row.names(simp_df)
        names(simp_smaller) <- c("average", "sd", "cumulative", "p", "mean_start", "mean_end")
        
        # save output to list
        simps[[i]] <- simp_smaller
        
        # if permanova wasn't significant, save a placeholder in the list   
    } else {
        simps[[i]] <- data.frame("None" = "SIMPER not run")
    }
    
    names(simps)[[i]] <- as.character(zn)
}
```


### Post-hoc output  

Was there a community difference between "Start" (first 2 years) and "End" (last 2 years) for each zone?  

```{r show-perm-posthoc, eval = datEnough}
pmvs_summ <- bind_rows(pmvs, .id = "Vegetation Zone") %>% 
    filter(term == "Time_group") %>% 
    select('Vegetation Zone', R2, p.value) %>% 
    mutate(Bonferroni_p.value = p.adjust(p.value, method = "bonferroni"))

kbl_nmst(pmvs_summ,
      digits = 4,
      caption = "Summary of pairwise PERMANOVAs testing community difference between first and most recent years of monitoring")
```



<details><summary>Click to see complete anova tables for each pairwise comparison</summary>  

Note - these individual tables do not contain Bonferroni-adjusted p-values; only the raw values.  

```{r show-pairwise-perm, results = 'asis', eval = datEnough}
for(i in seq_along(pmvs)){
    print(kbl_nmst(x = pmvs[[i]], 
                caption = paste0("Pairwise results for Vegetation Zone ", names(pmvs)[i]),
                digits = c(0, 0, 2, 4, 2, 4))
          )
}
```

</details>

## SIMPER  

```{r, results = "asis"}
# if there weren't enough years of data, can't do the rest.
# set up to NOT evaluate code chunks, and spit out a message.
datEnough <- nrow(dat_multivar) > 0

if(!datEnough){
    cat("### **Multivariate analyses were not run.**")
    cat("\n \nThere were not enough rows of data.")
}
```

If the PERMANOVA for a single zone had p-value < 0.2; *not* adjusted for multiple comparisons due to exploratory nature of these analyses, and note this does **not** represent a significant difference, a SIMPER was performed. **Please check the 'Post-hoc output' table in the PERMANOVA tab for the p-value, and be cautious in interpretation.** 

The top 6 species in output are below.  

"average" is the average contribution of that species to the Bray-Curtis distance between the two groups (note, this is *not* expressed in % and the column does not total to 1); "sd" is the standard deviation of the species' contribution. "cumulative" is the cumulative % contribution for this species and all those above it in the table. Typically people only report species up to the one that brings "cumulative" over 0.7. "p" is a p-value for that species based on permutation tests. "mean_start" is the mean cover of that species in the starting year(s), and "mean_end" is the mean cover of the species in the last year(s) of monitoring.  


```{r show-simper, results = 'asis', eval = datEnough}
for(i in seq_along(simps)){
    print(kbl_nmst(x = simps[[i]], 
                   caption = paste0(names(simps)[i], " SIMPER results; PERMANOVA p = ", round(pmvs_summ$p.value[i], 3)),
                   digits = 4)
    )
}
```

## NMDS  

```{r, results = "asis"}
# if there weren't enough years of data, can't do the rest.
# set up to NOT evaluate code chunks, and spit out a message.
datEnough <- nrow(dat_multivar) > 0

if(!datEnough){
    cat("### **Multivariate analyses were not run.**")
    cat("\n \nThere were not enough rows of data.")
}
```


```{r stats-run-nmds, eval = datEnough}
    mds_resp <- metaMDS(resp_pm, distance = "bray", k = 3,
                        autotransform = FALSE, trace = 0)

    data.scores <- as.data.frame(scores(mds_resp)$sites)
    species.scores <- as.data.frame(scores(mds_resp)$species)
    species.scores$species <- rownames(species.scores) 
    
    # only keep the species/groups listed as reserve choices, to reduce clutter on graphic
    # first get everything to group
    species.scores <- species.scores %>% 
        left_join(species_info, by = c("species" = "Species")) %>% 
        mutate(Species_or_Group = case_when(species %in% choices ~ species,
                                            Plant_Categories %in% choices ~ Plant_Categories,
                                            NMST_Groupings %in% choices ~ NMST_Groupings,
                                            Cover_Categories %in% choices ~ Cover_Categories,
                                            .default = "Other")) %>% 
        filter(Species_or_Group %in% choices) %>% 
        select(species = Species_or_Group,
               NMDS1,
               NMDS2,
               NMDS3)
    
    # then find centroids for the species-or-groups
    species.scores <- aggregate(cbind(NMDS1, NMDS2, NMDS3) ~ species, data = species.scores, FUN = mean)
    
    # data.scores$Vegetation_Zone <- expl$Vegetation_Zone
    # data.scores$Year <- expl$Year
    data.scores <- data.scores %>%
        mutate(Zone = expl_pm$Vegetation_Zone,
               Year = as.factor(expl_pm$Year),
               Time_group = as.factor(expl_pm$Time_group),
               Zone_Timegroup = as.factor(expl_pm$Zone_Timegroup)) %>% 
        separate(Zone, into = c("Zone_abbrev", "Zone_full"),
                 sep = "-")  
```

***  

Final 3-dimensional NMDS stress was **`r ifelse(datEnough, round(mds_resp$stress, 4), "NMDS not run")`**.  

#### 3-d NMDS plot  



```{r plot-nmds-3d, eval = datEnough, fig.width = 6, fig.height = 6}
# set up zone colors
zone_cols <- as.vector(palcols_zones)
names(zone_cols) <- NULL

# main plot - points for each zone
# (no centroids)
fig <- plot_ly() %>% 
    add_trace(data = data.scores,
              type = "scatter3d",
              mode = "markers",
              x = ~NMDS1,
              y = ~NMDS2,
              z = ~NMDS3,
              color = ~Zone_abbrev,
              colors = zone_cols,
              symbol = ~Time_group,
              symbols = c("circle", "square"),
              marker = list(size = 5,
                            opacity = 0.5)
    ) 

# now add lines, text, and cross-hatches for each identified loading
for (i in 1:nrow(species.scores)) {
    fig <- fig %>% 
        add_paths(
            x = c(0, species.scores$NMDS1[i]),
            y = c(0, species.scores$NMDS2[i]),
            z = c(0, species.scores$NMDS3[i]),
            # type = "scatter3d",
            mode = "lines",
            line = list(color = "red", width = 4),
            showlegend = FALSE
        )  |>
        add_text(
            x = species.scores$NMDS1[i],
            y = species.scores$NMDS2[i],
            z = species.scores$NMDS3[i],
            text = species.scores$species[i],
            textfont = list(color = "red", size = 12),
            showlegend = FALSE
        )  %>% 
        add_trace(
            type = "scatter3d",
            mode = "markers",
            x = species.scores$NMDS1[i],
            y = species.scores$NMDS2[i],
            z = species.scores$NMDS3[i],
            marker = list(size = 6,
                          color = "red",
                          symbol = "cross"),
            showlegend = FALSE
        )
        
}

# save the plot to another file
NMDS_outname <- paste0("NMDS-3d_", params$file_code, "_", Sys.Date(), ".html")
saveWidget(fig, file = here::here("output", NMDS_outname), selfcontained = TRUE)

# print the plot
fig
```

#### 2-dimensional NMDS plot  

(first two axes only)  

#### How to read these NMDS plots  

-  Large points represent the centroid for each Vegetation Zone/Time period combination. They are labelled with the Vegetation Zone abbreviation and 'Start' or 'End'. Additionally, they are colored by Vegetation Zone and shaped by 'Start' vs. 'End' time periods. Upon hover, the full Vegetation Zone name and year(s) represented will be provided.  
-  Black arrows and text represent Species or Categories of vegetation (e.g. H-Halophyte) identified by the reserve. Hovering over the arrow will provide the Species or Category name.  
-  Small points represent individual plots and are colored by Vegetation Zone and shaped by 'Start' vs. 'End' time periods (shapes and colors match the large points above); hovering does not yield output for points.  
-  Some identified loading factors may not appear as arrows; this is usually because all data points for that species were 0. They may also have been too low to have produced species scores in the NMDS.  


```{r stats-prep-nmds-groups, eval = datEnough}
# get centroids for each time/zone group
# https://stackoverflow.com/a/47523181
cent <- aggregate(cbind(NMDS1, NMDS2, NMDS3) ~ Zone_Timegroup, data = data.scores, FUN = mean)

en_coord_cat <- cent %>%
    separate(Zone_Timegroup, into = c("Zone", "Time_group"),
             sep = "; ") %>%
    left_join(keeps2,
              by = c("Zone" = "Vegetation_Zone",
                     "Time_group")) %>% 
    separate(Zone, into = c("Zone_abbrev", "Zone_full"),
             sep = "-") %>%
    mutate(Zone_Timegroup = paste(Zone_abbrev, Time_group),
           Zone_Years = paste(Zone_full, Years, sep = ": "))
```



```{r stats-graph-nmds-centroids, fig.width = 6, fig.height = 6, eval = datEnough}
p <- plot_nmds() 
girafe(ggobj = p,
       width_svg = 6,
       height_svg = 5,
       width = 0.8,
       height = 0.6
       )
# geom_text_repel is not yet implemented in plotly
# 
# p2 <- plot_nmds(axes = c(1, 3))
# girafe(ggobj = p2,
#        width_svg = 6,
#        height_svg = 5,
#        width = 0.8,
#        height = 0.6)
# 
# p3 <- plot_nmds(axes = c(2, 3))
# girafe(ggobj = p3,
#        width_svg = 6,
#        height_svg = 5,
#        width = 0.8,
#        height = 0.6)

```


