---
title: "Veg Analysis"
output: 
    html_document:
        toc: true
        toc_float: true
        toc_depth: 2
params: 
    file_code: "NIW"
---

<details>
<summary>General to-dos</summary>

To-dos:  

-  see sub-list in QAQC child Rmd  
-  show the list of reserve's choices (which species to highlight, which species represent ecotone invaders in each zone, etc.)    
-  make a script to install necessary packages, and a script to make sure packages are installed and loading  
-  general documentation on how a reserve can run this - will be a single script calling this .Rmd; actually should be easier for Reserves than SETr  

</details>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = TRUE)
```

```{r, message = FALSE, warning = FALSE}
# library(Hmisc)
library(emmeans)
library(multcomp)
library(multcompView)
library(lmerTest)
library(dplyr)
library(tidyr)
library(readxl)
library(stringr)
library(kableExtra)
library(DT)
library(ggplot2)


file_dat <- here::here("data", paste0(params$file_code, "_NMST.xlsx"))
file_specs <- here::here("metadata", paste0(params$file_code, "veg_specs.xlsx"))
```

```{r}
source(here::here("R", "sourced", "functions.R"))
```


```{r}
# read data, metadata, and analysis specifications
dat <- get_data(file_dat, cover_only = TRUE) # gets rid of density and height columns; keeps F_ columns
stn_tbl <- get_stn_table(file_dat)
species_info <- get_species_info(file_dat)
eis <- get_ecotone_invaders(file = file_specs)
anaSpecs <- get_anaSpecs(file = file_specs)
anaSpec_list <- split(anaSpecs, anaSpecs$R_anaName)

# add some columns to dat
dat <- dat %>% 
    dplyr::mutate(Date = lubridate::decimal_date(lubridate::ymd(paste(Year, Month, Day, sep = "-"))),
                  Years_sinceStart = round(Date - min(Date), 4),
                  StTrns = paste(SiteID, TransectID, sep = "-"),
                  StTrnsPlt = paste(SiteID, TransectID, PlotID, sep = "-")) %>% 
    dplyr::relocate(c(Date, Years_sinceStart), .before = Year) %>% 
    dplyr::relocate(c(StTrns, StTrnsPlt), .after = PlotID)
```


```{r}
# clean up data but keep some remnants for QA/QC tables
unsampds <- find_unsampleds(dat)
# find suspect data for later reporting
# NEED TO MAKE FLAG VALUES FLEXIBLE
susps <- find_suspect_values(dat, "-3")


dat <- remove_unsampleds(dat)  # get rid of rows full of NA
dat <- na_to_0(dat)  # turn remaining NAs into 0s
dat <- remove_suspect_values(dat, flags = c("-3"))  # remove suspect values
dat <- remove_unsampleds(dat)  # just in case a whole sample date was removed due to flagging - don't want a whole row of NAs



dat <- join_zones(dat, stn_tbl)  # add vegetation zones to each row
dat_long <- dat %>%
    pivot_longer(-(Reserve:Notes),
                 names_to = "Species",
                 values_to = "Cover")
```



# Exploration and QA/QC {.tabset}  

```{r, child = "sourced/001_QAQC.Rmd"}

```


# Summary Figures {.tabset}  

```{r, child = "sourced/002_Summary_Figures.Rmd"}

```


# Univariate Analyses {.tabset}  

```{r, child = "sourced/003_Univariate_Analyses.Rmd"}

```


# Multivariate Analyses {.tabset}  

```{r, child = "sourced/004_Multivariate_Analyses.Rmd"}

```


# Documentation   

<details><summary>R Session Info; click to expand</summary>

```{r}
sessionInfo()
```

</details>