---
title: "Veg Analysis"
output: 
    html_document:
        toc: true
        toc_float: true
        toc_depth: 2
params: 
    file_code: "GRB"
---

# About this Document  

On the left is a floating table of contents for navigating between major sections of output:  

-  Exploration and QA/QC  
-  Summary bar charts  
-  Univariate analyses  
-  Multivariate analyses  

Within each major section, content is split into **tabs** that are spread **horizontally** across the screen. By default, the *About* tab for each section is showing. Click on any tab to see specifics.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = TRUE)
```

**changed file paths here for new file naming after manual formatting**  

```{r pkgs-paths, message = FALSE, warning = FALSE}
# library(Hmisc)
library(emmeans)
library(multcomp)
library(multcompView)
library(lmerTest)
library(dplyr)
library(tidyr)
library(readxl)
library(stringr)
library(kableExtra)
library(DT)
library(ggplot2)
library(modelbased)  # for CIs of model predictions
library(see)  # for CIs of model predictions
library(viridis) # for colorblind-friendly viridis palette
library(vegan)


file_dat <- here::here("data", paste0(params$file_code, "_veg.xlsx"))
file_specs <- here::here("metadata", paste0(params$file_code, "_veg-specs.xlsx"))

options(knitr.kable.NA = '')
```

```{r source-fns}
source(here::here("R", "sourced", "functions.R"))
```


```{r import-data-meta}
# read data, metadata, and analysis specifications
dat <- get_data(file_dat, cover_only = TRUE) # gets rid of density and height columns; keeps F_ columns
stn_tbl <- get_stn_table(file_dat)
stn_tbl <- stn_tbl %>% 
    mutate(PlotID_full = paste(SiteID, TransectID, PlotID, sep = "-"))
species_info <- get_species_info(file_dat)
eis <- get_ecotone_invaders(file = file_specs)
anaSpecs <- get_anaSpecs(file = file_specs) |> 
    filter(!is.na(Choice))
anaSpec_list <- split(anaSpecs, anaSpecs$R_anaName)
zone_factor <- get_zone_ordering(file_specs)

# remove vegetation zones from eis data frame that don't appear in the station table
eis <- eis[eis$Vegetation_Zone %in% unique(stn_tbl$Vegetation_Zone), ]

# get zones in the same order in stn_tbl and eis, for later use
stn_tbl$Vegetation_Zone <- factor(stn_tbl$Vegetation_Zone, levels = zone_factor)
eis$Vegetation_Zone <- factor(eis$Vegetation_Zone, levels = zone_factor)

# add some columns to dat
dat <- dat %>% 
    dplyr::mutate(Date = lubridate::decimal_date(lubridate::ymd(paste(Year, Month, Day, sep = "-"))),
                  Years_sinceStart = round(Date - min(Date), 4),
                  StTrns = paste(SiteID, TransectID, sep = "-"),
                  StTrnsPlt = paste(SiteID, TransectID, PlotID, sep = "-")) %>% 
    dplyr::relocate(c(Date, Years_sinceStart), .before = Year) %>% 
    dplyr::relocate(c(StTrns, StTrnsPlt), .after = PlotID)
```


```{r mod-clean-data}
# clean up data but keep some remnants for QA/QC tables
unsampds <- find_unsampleds(dat)
# find suspect data for later reporting
# NEED TO MAKE FLAG VALUES FLEXIBLE
susps <- find_suspect_values(dat, "-3")


dat <- remove_suspect_values(dat, flags = c("-3"))  # remove suspect values. also removes F_columns.
dat <- remove_unsampleds(dat)  # should get rid of any dates where a plot was not sampled or was entirely rejected. 


dat <- na_to_0(dat)  # turn remaining NAs into 0s - this first makes sure all species columns are numeric


dat <- join_zones(dat, stn_tbl)  # add vegetation zones to each row
dat_long <- dat %>%
    pivot_longer(-(Reserve:Total),
                 names_to = "Species",
                 values_to = "Cover")
```

```{r mod-richness-diversity}
spps_live <- species_info$Species[species_info$Cover_Categories == "Live vegetation"]
# make sure they're in the data file too; several extras in NE files
spps_live <- spps_live[which(spps_live %in% names(dat))]

dat_live <- dat %>% 
    select(all_of(spps_live))
swdiv <- diversity(dat_live, index = "shannon")
rich <- specnumber(dat_live)

dat_div <- dat %>% 
    select(Reserve, StTrnsPlt, Year, Month, Day) %>% 
    mutate(SWdiv = swdiv,
           Richness = rich)

rm(dat_live, swdiv, rich)
```


```{r define-user-choices}
# generate a list of data frames based on user's analysis specs
anaSpec_dfs <- purrr::map(anaSpec_list, function(x) make_spec_df(dat_long, x$Choice))

# move unvegetated (if present) and other to end of species-as-factors
# for all anaSpec_dfs
for(i in 1:length(anaSpec_dfs)){
    tmp <- anaSpec_dfs[[i]]
    tmp$Species_or_Group <- relevel_spps(tmp$Species_or_Group)
    anaSpec_dfs[[i]] <- tmp
}
```


```{r mod-spp-summary}
# summarize species present
spp_summ <- 
    dat_long %>% 
    group_by(SiteID, Species) %>% 
    summarize(times_present = sum(Cover > 0),
              mean_cover = round(mean(Cover, na.rm = TRUE), 3),
              max_cover = round(max(Cover, na.rm = TRUE), 3)
              ) %>% 
    arrange(Species, SiteID)
```

```{r mod-attach-EIR}
# attach EIR to data frame
EIR <- dat_long %>% 
    left_join(eis) %>% 
    mutate(Invader = case_when(is.na(Invader) ~ 0,
                               .default = 1),
           EI_Cover = Cover * Invader) %>% 
    group_by(Year, Month, Day, Reserve, SiteID, TransectID, PlotID) %>% 
    summarize(EIR = round(sum(EI_Cover) / sum(Cover), 5))
# confirmed at April TWG that denominator is total cover

dat <- left_join(dat, EIR)
```

```{r define-eir-species}
eir_spps <- eis %>% 
    arrange(Vegetation_Zone, Species) %>% 
    mutate(Invader = case_when(Invader == 1 ~ "X",
                               .default = NA_character_)) %>% 
    pivot_wider(names_from = Vegetation_Zone,
                values_from = Invader) 
```


```{r mod-lump-spps, message = FALSE}
# combine species  

# top 3, for use in exploratory/spatial graphics
top_3_mean <- lump_species(dat_long, "mean", 3) %>% 
    dplyr::mutate(Date = lubridate::decimal_date(lubridate::ymd(paste(Year, Month, Day, sep = "-"))),
                  Years_sinceStart = round(Date - min(Date), 4),
                  PlotID_full = paste(SiteID, TransectID, PlotID, sep = "-"),
                  Species = relevel_spps(Species)) %>% 
    dplyr::relocate(c(PlotID_full, Date, Years_sinceStart), .before = Year)


top_3_wide <- top_3_mean %>% 
    pivot_wider(names_from = Species,
                values_from = Cover,
                values_fill = 0) %>% 
    group_by(PlotID_full) %>% 
    mutate()
```


```{r mod-group-nmst-cats}
dat_nmst_cats <- dat_long %>% 
    left_join(species_info) %>% 
    mutate(NMST_Groupings = case_when(is.na(NMST_Groupings) ~ Species,
                                      .default = NMST_Groupings)) %>% 
    select(-Species, -Plant_Categories, -Native_Classification, -Cover_Categories) %>% 
    pivot_wider(names_from = NMST_Groupings,
                values_from = Cover,
                values_fn = sum)
```


```{r define-plots-notEnoughReadings}
# find plots that don't have readings in at least 3 separate years
not_enough <- dat %>% 
    group_by(StTrnsPlt, Year) %>% 
    tally() %>% 
    ungroup() %>% 
    group_by(StTrnsPlt) %>% 
    summarize(n_years = length(Year)) %>% 
    filter(n_years < 3)
```


```{r define-color-pals}
# define color palettes - used in both QAQC and Summary Graphic child rmds

# top 3 species + other
palcols_top <- viridis(5, option = "magma")[1:4]  # don't use the yellow endpoint
names(palcols_top) <- levels(top_3_mean$Species)

# spatial barplots: up to 3 species/groups + other
n <- min(4, length(unique(anaSpec_dfs$spat$Species_or_Group)))
palcols_spec <- viridis(n)
names(palcols_spec) <- levels(anaSpec_dfs$spat$Species_or_Group)

# univariate exploration: up to 4 species/groups + other
n <- min(5, length(unique(anaSpec_dfs$univar$Species_or_Group)))
palcols_univar <- viridis(n + 1)[1:n]
names(palcols_univar) <- levels(anaSpec_dfs$univar$Species_or_Group)

# averaged stacked barplots: up to 8 species/groups + other
n <- min(9, length(unique(anaSpec_dfs$multivar$Species_or_Group)))
palcols_multivar <- viridis(n)
names(palcols_multivar) <- levels(anaSpec_dfs$multivar$Species_or_Group)


# also want zone colors to be consistent
n <- length(zone_factor)
palcols_zones <- viridis(n+1)[1:n]  # don't use the yellow endpoint
names(palcols_zones) <- levels(zone_factor)


# and site id
n <- length(unique(stn_tbl$SiteID))
palcols_sites <- viridis(n+1, option = "magma")[1:n]  # don't use the yellow endpoint
names(palcols_sites) <- unique(stn_tbl$SiteID)
```

# Exploration and QA/QC {.tabset}  

```{r source-001-QAQC, child = "sourced/001_QAQC.Rmd"}

```


# Summary Bar Charts {.tabset}  

```{r source-002-Figs, child = "sourced/002_Summary_Figures.Rmd"}

```


# Univariate Analyses {.tabset}  

```{r source-003-univarAna, child = "sourced/003_Univariate_Analyses.Rmd"}

```


# Multivariate Analyses {.tabset}  

```{r source-004-multivarAna, child = "sourced/004_Multivariate_Analyses.Rmd"}

```


# Documentation   

<details><summary>R Session Info; click to expand</summary>

```{r define-sessionInfo}
sessionInfo()
```

</details>