---
title: ''
output: html_document
---

```{r}
bySp <- bySp  %>%  
    dplyr::mutate(Date_centered = Date - median(Date),
                  Years_sinceStart = Date - min(Date),
                  Transect2 = paste(SiteID, TransectID, sep = "-"),
                  Plot2 = paste(SiteID, TransectID, PlotID, sep = "-"))

byGroup <- byGroup %>%  
    dplyr::mutate(Date = lubridate::decimal_date(lubridate::ymd(paste(Year, Month, Day, sep = "-"))),
                  Date_centered = Date - median(Date),
                  Years_sinceStart = Date - min(Date))
```

## Construct Models

### Linear Model (simple)  

```{r}
mod_lm <- lm(EIR ~ Years_sinceStart*Vegetation_Zone,
             data = bySp)
```


### Mixed Model, full nesting, lme4  

```{r}
mod_nested_lme4 <- lmer(EIR ~ Years_sinceStart*Vegetation_Zone + (1|SiteID/TransectID/PlotID),
                   data = bySp)
```

### Mixed Model, full nesting, nlme  

```{r}
mod_nested_nlme <- lme(EIR ~ Years_sinceStart*Vegetation_Zone, 
                       random = ~1|SiteID/TransectID/PlotID,
                       data = bySp,
                       na.action = na.omit)
```

### Mixed Model, Transect/Plot, lme4  

```{r}
mod_nested2_lme4 <- lmer(EIR ~ Years_sinceStart*Vegetation_Zone + (1|Transect2/PlotID),
                   data = bySp)
```

### Mixed Model, Transect/Plot, nlme  

```{r}
mod_nested2_nlme <- lme(EIR ~ Years_sinceStart*Vegetation_Zone, 
                       random = ~1|Transect2/PlotID,
                       data = bySp,
                       na.action = na.omit)
```

### Mixed Model, plots only, lme4  

```{r}
mod_plots_lme4 <- lmer(EIR ~ Years_sinceStart*Vegetation_Zone + (1|Plot2),
                   data = bySp)
```

### Mixed Model, plots only, nlme  

```{r}
mod_plots_nlme <- lme(EIR ~ Years_sinceStart*Vegetation_Zone, 
                       random = ~1|Plot2,
                       data = bySp,
                       na.action = na.omit)
```



Homemade function to get emtrends out of our models

```{r}
get_trends <- function(model){
    nm <- enexpr(model)
    tmp <- data.frame(emtrends(model, 
                      pairwise ~ Vegetation_Zone, 
                      var = "Years_sinceStart")$emtrends)
    # tmp$mod <- enexpr(model)
    # tmp$mod <- as.character(nm)
    return(tmp)
}
```

## Results  

### Models that didn't run  

this code definitely needs to be turned into a function or a loop  
preferable with informative error messages for the end user  
(they won't understand this terminology but for now it's what I'm trying to work out)  

```{r}
# allow for models that have errored and 
# add models that did work to a list to pull out results

mods <- list()


if(exists("mod_lm")){
    mods[["lm"]] <- mod_lm
} else {
    cat("mod_lm errored")
}

if(exists("mod_nested_lme4")){
    mods[["fully nested lme4"]] <- mod_nested_lme4
} else {
    cat("fully nested model from lme4 errored")
}

if(exists("mod_nested_nlme")){
    mods[["fully nested nlme"]] <- mod_nested_nlme
} else {
    cat("fully nested model from nlme errored")
}


if(exists("mod_nested2_lme4")){
    mods[["transect/plot lme4"]] <- mod_nested2_lme4
} else {
    cat("transect/plot model from lme4 errored")
}


if(exists("mod_nested2_nlme")){
    mods[["transect/plot nlme"]] <- mod_nested2_nlme
} else {
    cat("transect/plot model from nlme errored")
}


if(exists("mod_plots_lme4")){
    mods[["plots only lme4"]] <- mod_plots_lme4
} else {
    cat("plots only random effect model from lme4 errored")
}


if(exists("mod_plots_nlme")){
    mods[["plots only nlme"]] <- mod_plots_nlme
} else {
    cat("plots only random effect model from nlme errored")
}
```


```{r}
trends <- purrr::map(mods, function(x) get_trends(x))
trends_df <- bind_rows(trends, .id = "model") %>% 
    mutate(model = forcats::as_factor(model))
```

Pull out marginal and conditional R^2 for mixed models and add columns to df  

```{r, warning = FALSE, message = FALSE}
rsqds <- purrr::map(mods, function(x) as.data.frame(MuMIn::r.squaredGLMM(x)))
rsqds_df <- bind_rows(rsqds, .id = "model") %>% 
    mutate(model = forcats::as_factor(model))
trends_df <- left_join(trends_df, rsqds_df)
```



### Singular Models  

```{r}
check_singularity <- function(model){
    # model needs to be from lme4
    nm <- enexpr(model)
    mod <- model
    cat(paste0(nm, ": \n"))
    if(isSingular(mod)){
        cat("is singular. VarCorr follows. \n \n")
        print(summary(mod)$varcor)
        cat("\n \n")
    } else {
        cat("fit is okay (not singular) \n \n")
    }
}
```


```{r}
if(exists("mod_nested_lme4")){check_singularity(mod_nested_lme4)}
if(exists("mod_nested2_lme4")){check_singularity(mod_nested2_lme4)}
if(exists("mod_plots_lme4")){check_singularity(mod_plots_lme4)}
```

### Table  


```{r}
to_print <- trends_df %>% 
    mutate(across(Years_sinceStart.trend:R2c, function(x) round(x, 4)),
           df = round(df, 2)) %>% 
    arrange(Vegetation_Zone, model) %>% 
    rename(TrendPerYear = Years_sinceStart.trend) %>% 
    relocate(Vegetation_Zone)
    
kbl(to_print) %>% 
    kable_paper(full_width = F) %>%
    column_spec(1, bold = T) %>%
    collapse_rows(columns = 1, valign = "top")
```


### Numerical effects  

For the model with transect/plot nesting.

```{r}
if(exists("mod_nested2_lme4")){
    print(anova(mod_nested2_lme4))
    mod <- lmerTest::as_lmerModLmerTest(mod_nested2_lme4)   
    cat("\n \n Above is anova table of lme4 output; \n Below is anova table from lmerTest. \n \n ")
    
    mod_anov <- anova(mod, ddf = "Kenward-Roger")
    tidied <- broom::tidy(mod_anov)
    sig_int <- tidied[tidied$term == "Years_sinceStart:Vegetation_Zone", "p.value"] < 0.05
    print(mod_anov)
    cat("\n \n")
    emtrends(mod_nested2_lme4, 
                  pairwise ~ Vegetation_Zone, 
                  var = "Years_sinceStart")
}
```


## Graphs  

```{r, warning = FALSE, message = FALSE}
if(exists("mod_nested2_lme4")){
    pdf <- emmip(mod_nested2_lme4, Vegetation_Zone ~ Years_sinceStart,
                 cov.reduce = range,
                 linearg = list(size = 1), CIs = TRUE,
                 plotit = FALSE)
    
    p1 <- ggplot() +
        geom_line(data = pdf, aes(x = Years_sinceStart,
                                  y = yvar,
                                  col = Vegetation_Zone,
                                  linetype = Vegetation_Zone),
                  linewidth = 0.8) +
        theme_bw()
    
    
    
    p2 <- ggplot() +
        geom_jitter(data = bySp, aes(x = Years_sinceStart, y = EIR,
                                     col = Vegetation_Zone),
                    alpha = 0.7) +
        geom_line(data = pdf, aes(x = Years_sinceStart,
                                  y = yvar,
                                  col = Vegetation_Zone,
                                  linetype = Vegetation_Zone),
                  linewidth = 0.8) +
        facet_wrap(~Vegetation_Zone) +
        theme_bw() +
        theme(legend.position = "none")
    
    print(p1 + ggtitle("Mixed Model fit"))
    print(p2)
}
```


### Contrasts plot  

For the model with transect/plot nesting. If the interaction between vegetation zone and time was significant (p < 0.05), letters are used to represent groups of slopes that are not significantly different (via pairwise comparisons). If the interaction was not significant, this plot represents the marginal slopes but no letters are used.

Some notes from the function. Kenward-Roger method used to estimate denominator degrees of freedom. Confidence level used: 0.95. p-value adjustment: Tukey. alpha = 0.05.  

"If two or more means share the same grouping symbol,
      then we cannot show them to be different.
      But we also did not show them to be the same."

```{r}
if(exists("mod_nested2_lme4")){
    trnds <- emtrends(mod_nested2_lme4, 
                  pairwise ~ Vegetation_Zone, 
                  var = "Years_sinceStart")
    mod_cld <- cld(trnds, Letters = letters) %>% 
        mutate(.group = str_trim(.group))
    
    p <- ggplot(mod_cld, aes(x = reorder(Vegetation_Zone, Years_sinceStart.trend),
                             y = Years_sinceStart.trend,
                             label = .group,
                             color = Vegetation_Zone)) +
        geom_pointrange(aes(ymin = lower.CL, 
                            ymax = upper.CL),
                        linewidth = 0.8) +
        labs(title = "Rate of change in EIR by zone",
             x = "Vegetation Zone",
             y = "change per year") +
        theme_bw() +
        theme(legend.position = "none")
    
    if(sig_int){
        p <- p +
            geom_text(aes(y = max(mod_cld$upper.CL)),
                      nudge_y = max(mod_cld$upper.CL)/10,
                      color = "black")
    }
    print(p)
}
```

