---
title: "Playing with permutations"
format: html
---

Possibly run separately for each site
possibly use how() with SiteID as block, TransectID as “plot” and within = “free” - would allow transects to be exchanged within (but not between) a site, and plots to be exchanged within a transect.


```{r}
library(tidyverse)
library(vegan)
```

# Make a toy dataset  

Needs multiple levels: SiteID, TransectID, PlotID. And "readings" across multiple "years".  3 sites, each with 3 transects, each with 3 plots, each read in 3 years

```{r}
yrs <- rep(2010:2012, each = 27)
sitesPerYear <- rep(1:3, each = 9)
sites <- rep(sitesPerYear, 3)
transectsPerSite <- rep(1:3, each = 3)
transects <- rep(transectsPerSite, 9)
plots <- rep(1:3, 27)
set.seed(2023)


dat <- data.frame("Year" = yrs,
                 "SiteID" = sites,
                 "TransectID" = transects,
                 "PlotID" = plots,
                 "Reading" = 1:81)
```


# Permute  

## unrestricted

```{r}
perm <- shuffleSet(dat)
```


```{r}
dat2 <- dat %>% arrange(SiteID, TransectID)
```


```{r}
shuffle(dat2, how(within = Within(type = "free"),
                 plots = Plots(strata = factor(dat2$TransectID)),
                 blocks = factor(dat2$SiteID)))
```


```{r}
toy <- data.frame("Letters" = c("A", "B", "C", "D", "E"),
                  "Numbers" = 1:5)

toy
shuffleSet(toy)
shuffle(toy)

```

